---
title: "Compare biomass between years"
output: html_notebook
---

# Objective of code
Analysis 1: Are there differences in acosutically detected biomass and CM between years?! 

# Background: 
The following code analyzes whether there were any differences in the biomass and center of mass depth (m) BETWEEN years in our study. We are interested in inter annual differences AND at varying cell sizes (sensitivity analysis)

NOTE: 
- analysis part 1 (this document): Comparing differences in biomass and Center of mass depth between years (and diff spatial scales of regridding)
- analysis part 2 (Analysis 2.Rmd): which oceanographic variables are likely driving the center of mass depth of the DSL at a regridded distance of 25km


# Load libraries 
```{r load libraries}
library(tidyverse) # all things
library(lubridate) # dates
library(sf) # spatial features, mapping
library(multcompView) #cld compact letter display for output
library(rcompanion) # this just helps us complete the mult comp test output so it can be read by mulcompView
library(ggthemes) # simplification of themes for ggplotting
```

# Which re-grid size are you working on?!! MANUAL step
I regridded by 1km, 5km, 10km, 15km, 20km (from fine-resolution, 100m data originally output)
Need to manually set the dataset here (df are differentiated by grid size as acoustics_regrid_15km for ex) so the following code will all run on the same input 
```{r manually set the grid size we are working with }

# MANUAL step: select which regridding of acoustics data we are interested in
acoustic_regrid= acoustics_regrid_25km

grid_size="25km" #MANUAL step

```

### MINIMUM number of 100m grid cells to trust in larger grid size
There are some instances where our re-gridded cell (of whichever scale) were based on 1 or few 100m horizontal cells. Since we are extrapolating these values over a large region, select those cells with at least a minimum of 3 100m fine-resolution cells covering a larger grid cell

```{r}
# Ensure each cell has at least 3 100m grid cells per re-gridded cell (echo_count)

acoustic_regrid= acoustic_regrid |> 
                 filter(echo_count>=3)

```

# set up gridlines (regardless of regrid size) for plotting

```{r}

# gridlines-- to draw gridlines based on indivdiual gridcell size
#longitude (x axis)
lon1 <- unique(acoustic_regrid$lon_grid_low)
lon2<- unique(acoustic_regrid$lon_grid_hi)
# add on highest value from lon_grid_high
lon=unique(c(lon1, lon2))
lon
rm(lon1, lon2)

#latitude (y axis)
lat1 <- unique(acoustic_regrid$lat_grid_low)
lat2<- unique(acoustic_regrid$lat_grid_hi)
# add on highest value from lon_grid_high
lat=unique(c(lat1, lat2))
lat
rm(lat1, lat2)


```

# PLOT acoustic track over regridded data-- how many fine-resolution cells per re-gridded cell?
need to adjust the shape size depending on grid size ~4 is a nice size for 25km

#MANUAL: change cell size in geom_point
```{r}

# MANUAL: change the size in geom_point to fit grid cell size :)
ggplot()+
  geom_sf(data=states, fill="grey", color="black")+# plot state basemap
  geom_point(data = acoustic_regrid, aes(x= lon, y=lat, fill=MVBS), color="black", stroke=0, size=5.2, shape=22 )+ #plot regridded data
  geom_text(data=acoustic_regrid, aes(x=lon, y=lat, label=echo_count), size=2)+ #label # of 100m cells per re-grid cell
  geom_point(data=echo_integration_all, aes(x=Lon_M, y=Lat_M), size=0.01, color="grey", alpha=0.9)+ #overlay of cruise track over regridded data
  scale_fill_viridis_c(option = "D", direction=1)+ #color output via viridis
  coord_sf(xlim = c(min(acoustic_regrid$lon-0.25), max(acoustic_regrid$lon+0.25)), ylim = c(min(acoustic_regrid$lat-0.25), max(acoustic_regrid$lat+0.25)))+ 
  scale_x_continuous(breaks = lon, minor_breaks = NULL)+
  scale_y_continuous(breaks=lat, minor_breaks = NULL)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid = element_line(colour = "black", linetype = "solid", linewidth =0.25),
        panel.background = element_blank())+
  labs(x="Longitude", y="Latitude", color = "MVBS")+ #legend
  facet_wrap(~year)
  
  
#Save plot
ggsave(filename=paste0("./plots/analysis 1/regrid_coverage_", grid_size, "_map.pdf"))
```
25km: size=5.2

# Calculate median values for plotting 
se=s/sqrt(n) where s= sample standard deviation and n=sample size. in our case, our units here are the # of grid cells per year
```{r}

# group by year and create summary stats such as median values and se and other labels for boxplot
summary_year <- acoustic_regrid  |> 
                group_by(year) |> 
                summarize(no_grid_cells=n(),
                          median_MVBS=round(median(MVBS), 2),
                          #median_NASC=round(median(NASC), 5), 
                          median_CM=round(median(cm_m)),
                          se_MVBS=sd(MVBS)/sqrt(no_grid_cells),
                          #se_NASC=sd(NASC)/sqrt(no_grid_cells),
                          se_CM=sd(cm_m)/sqrt(no_grid_cells),
                          sd_MVBS=sd(MVBS),
                          #sd_NASC=sd(NASC),
                          sd_CM=sd(cm_m),
                          quantile_mvbs=quantile(MVBS, probs=0.75), #find the 3rd quantile for labeling letters later
                          quantile_cm=quantile(cm_m, probs=0.75)) 
```
I loaded the library plotrix which has the function std.error as a double check and all values the same so just going with sd()/sqrt(# grid cells)

# MVBS between years (boxplot)
The following is a box plot (with points shown) depicting MVBS values per year for our given grid cell size defined by 
```{r}

# simple box plot of MVBS values per year
ggplot(data=acoustic_regrid, aes(x=year, y=MVBS), color="black")+
  stat_boxplot(geom="errorbar", width=0.25)+
  geom_boxplot(data=acoustic_regrid, aes(x=year, y=MVBS, color=year), outlier.shape=NA, show.legend = FALSE)+ #plot boxplots without outliers since these are all plotted via geom_jitter
  scale_color_manual(values = c("2013"="grey20", "2014"="grey20", "2015"="chocolate", "2016"="chocolate", "2017"="grey20", "2018"="grey20"))+
  geom_jitter(alpha=0.2, color="snow4")+ #add actual data points to our boxplot
  geom_text(data= summary_year, aes(x = as.factor(year), y = median_MVBS, label = paste0("MVBS: ", median_MVBS)), size = 2.5, vjust = -1)+ #label median 
  geom_text(data= summary_year, aes(x = as.factor(year), y = median_MVBS, label = paste0("n: ", no_grid_cells)), size = 2, vjust = 1.5,color="black")+
  xlab("Year")+
  ylab("MVBS (dB)")+
  ggthemes::theme_few()


#save
ggsave(filename=paste0("./plots/analysis 1/boxplot_mvbs_", grid_size, ".pdf"), width=7, height=5, units="in")

```

# CM between years (boxplot)
```{r}
# simple box plot of NASC values
ggplot(data=acoustic_regrid, aes(x=year, y=cm_m, color=year))+
  stat_boxplot(geom="errorbar", width=0.25)+
  geom_boxplot(data=acoustic_regrid, aes(x=year, y=cm_m), outlier.shape=NA, show.legend = FALSE)+ #plot boxplots without outliers since in jitter
  scale_color_manual(values = c("2013"="grey20", "2014"="grey20", "2015"="chocolate", "2016"="chocolate", "2017"="grey20", "2018"="grey20"))+
  geom_jitter(alpha=0.1, color="slategrey")+ #add actual data points to our boxplot
  scale_y_reverse()+
  geom_text(data= summary_year, aes(x = as.factor(year), y = median_CM, label = paste0("CM: ", median_CM)), size = 2.5, vjust = -0.5)+ #label median 
  geom_text(data= summary_year, aes(x = as.factor(year), y = median_CM, label = paste0("n: ", no_grid_cells)), size = 2, vjust = 1.5, color="black")+ #label sample size 
  xlab("Year")+
  ylab("CM (m)")+
  ggthemes::theme_few()+
  theme(legend.position = 'none')

# save plot to plots folder
ggsave(filename=paste0("./plots/analysis 1/boxplot_cm_", grid_size, ".pdf"), width=7, height=5, units="in")
```
# PLOT: Distribution of MVBS between years
NOTE:using colors from Okabe & Ito (2008) 
```{r}
# using binned output (25x25km)
ggplot(data = acoustic_regrid, aes(x=MVBS, colour=year))+
  #geom_histogram(aes(y=..density..),color="grey", fill="transparent", binwidth = 50)+ #this is plotting properly for some reason
  geom_density(alpha=0.2, fill="NA", linewidth=1)+ #plot density curves
  scale_color_manual(values = c("2013"="#009E73" , "2014"= "#56B4E9", "2015"="#E69F00", "2016"="#D55E00" , "2017"= "#0072B2", "2018"= "#999999"))+
  scale_x_continuous(expand=c(0,0), breaks = round(seq(min(acoustic_regrid$MVBS), max(acoustic_regrid$MVBS), by=1), 0))+ #move plot to axis
  scale_y_continuous(expand = c(0,0))+
  xlab("MVBS (dB) ")+
  ylab("Density")+
  labs(color="Year")+
  theme_classic()+
  theme(axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10))

ggsave(paste0("./plots/analysis 1/mvbs_bw_years_", grid_size, ".pdf"))
```


Great little plot depicting the differences in the distribution of MVBS between years! 

# Plot the distribution of CM between years
```{r}
# using binned output (25x25km)

ggplot(data = acoustic_regrid, aes(x=cm_m, colour=year))+
  geom_density(alpha=0.2, fill="NA", linewidth=1)+ #plot density curves
  scale_color_manual(values = c("2013"="#009E73" , "2014"= "#56B4E9", "2015"="#E69F00", "2016"="#D55E00" , "2017"= "#0072B2", "2018"= "#999999"))+
  scale_x_continuous(expand=c(0,0), breaks = round(seq(min(acoustic_regrid$cm_m)-4, max(acoustic_regrid$cm_m)+1, by=25), 0))+ #move plot to axis
  scale_y_continuous(expand = c(0,0))+
  xlab("Center of mass (m) ")+
  ylab("Density")+
  labs(color="Year")+
  theme_classic()+
  theme(axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10))

ggsave(paste0("./plots/analysis 1/cm_bw_years_", grid_size, ".pdf"))
```

# Combined plot
The following is simply the two plots above combined into one output...
```{r}
library(patchwork)
p1=ggplot(data = acoustic_regrid, aes(x=MVBS, colour=year))+
  #geom_histogram(aes(y=..density..),color="grey", fill="transparent", binwidth = 50)+ #this is plotting properly for some reason
  geom_density(alpha=0.2, fill="NA", linewidth=1, show.legend = FALSE)+ #plot density curves
  scale_color_manual(values = c("2013"="#009E73" , "2014"= "#56B4E9", "2015"="#E69F00", "2016"="#D55E00" , "2017"= "#0072B2", "2018"= "#999999"))+
  scale_x_continuous(expand=c(0,0), breaks = round(seq(min(acoustic_regrid$MVBS), max(acoustic_regrid$MVBS), by=2), 0))+ #move plot to axis
  scale_y_continuous(expand = c(0,0))+
  xlab("MVBS (dB) ")+
  ylab("Density")+
  labs(color="Year")+
  theme_classic()+
  theme(axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10))
p2=ggplot(data = acoustic_regrid, aes(x=cm_m, colour=year))+
  geom_density(alpha=0.2, fill="NA", linewidth=1)+ #plot density curves
  scale_color_manual(values = c("2013"="#009E73" , "2014"= "#56B4E9", "2015"="#E69F00", "2016"="#D55E00" , "2017"= "#0072B2", "2018"= "#999999"))+
  scale_x_continuous(expand=c(0,0), breaks = round(seq(min(acoustic_regrid$cm_m)-4, max(acoustic_regrid$cm_m)+1, by=25), 0))+ #move plot to axis
  scale_y_continuous(expand = c(0,0))+
  xlab("Center of mass (m) ")+
  ylab("Density")+
  labs(color="Year")+
  theme_classic()+
  theme(axis.text.x = element_text(size=10),
        axis.text.y = element_text(size=10))
p1+p2
ggsave(paste0("./plots/analysis 1/cm_mvbs_dist_combo_", grid_size, ".pdf"), height=5, width=10, units = "in", dpi = 300)
```
# EXPORT plot with black background
```{r}
library(patchwork)

# MVBS plot
p1=ggplot(data = acoustic_regrid, aes(x=MVBS, colour=year))+
  #geom_histogram(aes(y=..density..),color="grey", fill="transparent", binwidth = 50)+ #this is plotting properly for some reason
  geom_density(alpha=0.2, fill="NA", linewidth=1, show.legend = FALSE)+ #plot density curves
  scale_color_manual(values = c("2013"="#009E73" , "2014"= "#56B4E9", "2015"="#E69F00", "2016"="#D55E00" , "2017"= "#0072B2", "2018"= "#999999"))+
  scale_x_continuous(expand=c(0,0), breaks = round(seq(min(acoustic_regrid$MVBS), max(acoustic_regrid$MVBS), by=2), 0))+ #move plot to axis
  scale_y_continuous(expand = c(0,0))+
  xlab("MVBS (dB) ")+
  ylab("Density")+
  labs(color="Year")+
  theme_classic()+
  theme(axis.text.x = element_text(size=10, color="white"),
        axis.text.y = element_text(size=10, color="white"),
        panel.background = element_rect(fill = 'black', color="white"),
        plot.background = element_rect(fill='black', color="white"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.background = element_rect(fill='transparent'),
        legend.box.background=element_rect(fill = 'transparent'),
        legend.text = element_text(color = "white"),
        legend.title = element_text(color="white"),
        axis.line = element_line(color = "white"),
        axis.title = element_text(color="white"))


# center of mass plot
p2=ggplot(data = acoustic_regrid, aes(x=cm_m, colour=year))+
  geom_density(alpha=0.2, fill="NA", linewidth=1)+ #plot density curves
  scale_color_manual(values = c("2013"="#009E73" , "2014"= "#56B4E9", "2015"="#E69F00", "2016"="#D55E00" , "2017"= "#0072B2", "2018"= "#999999"))+
  scale_x_continuous(expand=c(0,0), breaks = round(seq(min(acoustic_regrid$cm_m)-4, max(acoustic_regrid$cm_m)+1, by=25), 0))+ 
  scale_y_continuous(expand = c(0,0))+
  xlab("Center of mass (m) ")+
  ylab("Density")+
  labs(color="Year")+
  theme_classic()+
  theme(axis.text.x = element_text(size=10, color="white"),
        axis.text.y = element_text(size=10, color="white"),
        panel.background = element_rect(fill = 'black', color="white"),
        plot.background = element_rect(fill='black', color="white"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.background = element_rect(fill='transparent'),
        legend.box.background=element_rect(fill = 'transparent'),
        legend.text = element_text(color = "white"),
        legend.title = element_text(color="white"),
        axis.line = element_line(color = "white"),
        axis.title = element_text(color="white"))

p1+p2
ggsave(paste0("./plots/analysis 1/cm_mvbs_dist_combo_", grid_size, "_clear.pdf"), height=5, width=10, units = "in", dpi = 300, bg="transparent")
```


# STATISTICAL TEST:  KRUSKAL WALLIS
From Conover 1999: The following Assumptions are required:
1. All samples are random samples from their respective populations.
2. In addition to independence within each sample, there is mutual independence among the various samples.
3. The measurement scale is at least ordinal
4. Either the k population distribution functions are identical, or else some of the populations tend to yield larger values than other populations do. 

Hypotheses:
H0: all of the K population distribution functions are identical
H1: At least one of the populations tends to yield larger observations than at least one of the other populations
alt H1: The k populations do not all have identical means

# convert year to a factor
Not sure if this is patently required, but helps simplify structure 
```{r}
# first ensure that year is a factor
acoustic_regrid$year <- as.factor(acoustic_regrid$year)
levels(acoustic_regrid$year) #check on order of factor year (all good in sequential order)
```

# Kruskall Wallis test: MVBS and CM between years 
# MANUAL step
Make sure to change the grid cell size when storing output
```{r MANUAL STEP}

# MANUAL OUTPUT -- save output for grid size

# Kruskal wallis test-- MVBS
kw_mvbs= kruskal.test(MVBS ~ year, data = acoustic_regrid)
kw_mvbs

#MANUAL: save output per grid size
kw_mvbs_25km=kw_mvbs


# Kruskal wallis test--- CM
kw_cm= kruskal.test(cm_m ~ year, data = acoustic_regrid)
kw_cm

#MANUAL: save output per grid size
kw_cm_25km=kw_cm

```
If p-value <0.05 we can conclude that there are significant differences between groups (years). In order to know which pairs of years were different, we must further test with multiple comparisons.

# MUTLIPLE COMPARISONS-- Kruskal Wallis 
Although we know there are differences between years, we don't know between WHICH years these differences occur. To compare between groups (years) we thus additionally make pairwise comparisons.

Background: "If, and only if the null hypothesis is rejected, we may we use the following procedure to determine which pairs of populations tend to differ" uses rank-sum Conover

NOTE: default value for multiple comparisons is "BH" for Bejamini & Hochberg. see p.adjust for more info. Bonferonni is the most conservative option

#MANUAL
```{r mutliple comparisons kruskal wallis }
# multiple comparisons test-- MVBS
kw_mc_mvbs= pairwise.wilcox.test(acoustic_regrid$MVBS, acoustic_regrid$year,
                 p.adjust.method = "bonferroni")
kw_mc_mvbs

# MANUAL: save output per grid cell size
kw_mc_mvbs_25km= kw_mc_mvbs



# multiple comparisons test-- CM
kw_mc_cm= pairwise.wilcox.test(acoustic_regrid$cm_m, acoustic_regrid$year,
                 p.adjust.method = "bonferroni")

# MANUAL: save output per grid cell size
kw_mc_cm_25km= kw_mc_cm

# to see p-values:
format(kw_mc_cm$p.value, scientific = FALSE)


```

# Data for compact letter display (cld) via {rcompanion} to complete output table and {multcompView} for creating the CLD
compact lettering display
Compact letter display allows us to appropraitely color signifiant differences in our data
```{r}

#Take our p.value output from Pairwise.Wilcoxon.Text (ie mutliple comparisons for kruskall wallis)

kw_mc_cm
kw_mc_mvbs

# select for our p-value output- table of P-values for pairwise comparisons
table_cm= kw_mc_cm$p.value
table_mvbs= kw_mc_mvbs$p.value

#convert these tables into "full" tables ie not just NA using package rcompaniion
kw_mc_cm_full = rcompanion::fullPTable(table_cm)
kw_mc_mvbs_full = rcompanion::fullPTable(table_mvbs)


# calculate CLD
cld_cm= multcompView::multcompLetters(kw_mc_cm_full)
cld_mvbs= multcompView::multcompLetters(kw_mc_mvbs_full)

# display groupings
cld_cm
cld_mvbs


# add these values to our summary_year table
summary_year= summary_year |> 
  mutate(cld_cm=as.vector(cld_cm$Letters), cld_mvbs=as.vector(cld_mvbs$Letters))

#remove intermediary steps
rm(kw_mc_cm_full, kw_mc_mvbs_full, table_cm, table_mvbs)
```
NOTE: error message stating atomic vector error, but correctly added to df: summary_year so okay to proceed

# Add output from multiple comparisons to boxplot

see https://r-graph-gallery.com/84-tukey-test a little complicated
https://statdoe.com/one-way-anova-and-box-plot-in-r/

# Add multipe comparisons letters to boxplot -- MVBS
color blind friendly palette

#Updated
clrs = c(
"a" = "#009E73",
"b" = "#56B4E9",
"c" = "#E69F00",
"d" = "#D55E00",
"e" = "#0072B2",
"f" = "#999999"
)
# color palette from Okabe-Ito 

#MANUAL: adjust color ramp and position of letters within circles :) 

```{r}

#set up MANUAL color scale for letters
clrs= unique(summary_year$cld_mvbs)
clrs #run this to see what the output is for a given grid size
clrs = c(
"a" = "#0072B2",
"b" = "#E69F00"
)
# manually set some color-blind friendly colors and manually update labels depending on unique combinations

# simple box plot of MVBS values- multiple comparison results from KW 
ggplot(data=acoustic_regrid, aes(x=year, y=MVBS))+
  stat_boxplot(geom="errorbar", width=0.25)+ #error bars
  geom_boxplot(data=acoustic_regrid, aes(x=year, y=MVBS), outlier.shape=NA)+ #plot boxplots without outliers since in jitter
  geom_jitter( width=0.15, height=0, alpha=0.3, color="#999999" )+ #add actual data points to our boxplot don't jitter in vertical direction and close toge horiz
  geom_text(data= summary_year, aes(x = year, y = median_MVBS, label = paste0("MVBS: ", median_MVBS)), size = 2.8, vjust = -0.5)+ #label median 
  geom_text(data= summary_year, aes(x = year, y = median_MVBS, label = paste0("n: ", no_grid_cells)), size = 2.5, vjust = 1.5, color="black")+ #n
  geom_point(data = summary_year, aes(x = year, y = quantile_mvbs+1, label = cld_mvbs, color=cld_mvbs), size=6, alpha=1, show.legend = FALSE)+ # label cld
  geom_text(data = summary_year, aes(x = year, y = quantile_mvbs, label = cld_mvbs), vjust=-2.2, color="black")+ # label significance cl adjust position
  scale_color_manual(values=clrs)+ #use manual color palette defined above 
  xlab("Year")+
  ylab("MVBS (dB)")+
  theme(legend.position = 'none')+
  ggthemes::theme_few()


ggsave(filename=paste0("./plots/analysis 1/boxplot_kw_mc_mvbs_jitter_", grid_size, "_.pdf"), width=7, height=5, units="in")

```
Play around with the position of letters and circles, they shift with each cell size (eye roll)
Also vary when exported
25km: vjust= -2.2
20km: vjust= -1.7
15km: vjust= -1.4
10km: vjust= -1.3
5km: vjust= -1.4
1km: vjust= -1.3

# Same plot but transparent background
for presentations with a black background
```{r}
ggplot(data=acoustic_regrid, aes(x=year, y=MVBS))+
  stat_boxplot(geom="errorbar", width=0.25, color="white")+
  geom_boxplot(data=acoustic_regrid, aes(x=year, y=MVBS), fill=NA, color="white", outlier.shape=NA)+ #plot boxplots without outliers since in jitter
  geom_jitter( width=0.15, height=0, alpha=0.3, color="#999999" )+ #add actual data points to our boxplot don't jitter in vertical
  geom_text(data= summary_year, aes(x = year, y = median_MVBS, label = paste0("MVBS: ", median_MVBS)), size = 2.8, vjust = -0.5, color="white")+ #label median 
  geom_text(data= summary_year, aes(x = year, y = median_MVBS, label = paste0("n: ", no_grid_cells)), size = 2.5, vjust = 1.5, color="white")+ #n
  geom_point(data = summary_year, aes(x = year, y = quantile_mvbs+1, label = cld_mvbs, color=cld_mvbs), size=6, alpha=1, show.legend = FALSE)+ # label cld
  geom_text(data = summary_year, aes(x = year, y = quantile_mvbs, label = cld_mvbs), vjust=-2.2, color="white")+ # label significance cl
  scale_color_manual(values=clrs)+ #use manual color palette defined above 
  xlab("Year")+
  ylab("MVBS (dB)")+
  theme(legend.position = 'none')+
  ggthemes::theme_few()+
  theme(legend.position = 'none',
        panel.background = element_rect(fill = 'transparent', color="white"),
        plot.background = element_rect(fill='transparent', color="white"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.background = element_rect(fill='transparent'),
        legend.box.background=element_rect(fill = 'transparent'),
        axis.text = element_text(colour="white"),
        axis.line = element_line(color = "white"),
        axis.title = element_text(color="white"))


ggsave(filename=paste0("./plots/analysis 1/boxplot_kw_mc_mvbs_transparent_", grid_size, "_.pdf"), width=7, height=5, units="in", bg="transparent")
# Add multiple comparisons to boxplot -- CM
```

# # Add multipe comparisons letters to boxplot -- CM
```{r}

#set up MANUAL color scale for letters (manually change inputs to match cld output)
clrs= unique(summary_year$cld_cm)
clrs # run this to see unique output per grid size
clrs = c(
"a" = "#009E73",
"b" = "#56B4E9",
"c" = "#E69F00"
)

# manually set some color-blind freindly colors and manually update labels depending on unique combinations

# simple box plot of MVBS values- multiple comparison results from KW 
ggplot(data=acoustic_regrid, aes(x=year, y=cm_m, color=year))+
  stat_boxplot(geom="errorbar", width=0.25)+
  geom_boxplot(data=acoustic_regrid, aes(x=year, y=cm_m), outlier.shape=NA)+ #plot boxplots without outliers since in jitter
  #scale_color_manual(values = c("2013"="grey20", "2014"="grey20", "2015"="chocolate", "2016"="chocolate", "2017"="grey20", "2018"="grey20"))+
  geom_jitter(width=0.15, height=0, alpha=0.3, color="#999999" )+ #add actual data points to our boxplot
  geom_text(data= summary_year, aes(x = year, y = median_CM, label = paste0("CM: ", median_CM)), size = 2.8, vjust = -0.5, color="black")+ #label median 
  geom_text(data= summary_year, aes(x = year, y = median_CM, label = paste0("n: ", no_grid_cells)), size = 2.5, vjust = 1.5, color="black")+ #n
  geom_point(data = summary_year, aes(x = year, y = quantile_cm+10, label = cld_cm, color=cld_cm), size=6, alpha=1)+ # label cld
  geom_text(data = summary_year, aes(x = year, y = quantile_cm, label = cld_cm), vjust=1.8, color="black")+ # label significance cld
  scale_color_manual(values=clrs)+ #use manual color palette defined above 
  scale_y_reverse()+
  xlab("Year")+
  ylab("CM (m)")+
  ggthemes::theme_few()+
  theme(legend.position = 'none')
  
ggsave(filename=paste0("./plots/analysis 1/boxplot_kw_mc_cm_jitter_", grid_size, "_.pdf"), width=7, height=5, units="in", bg="transparent")

```
25km: vjust= 2.1
20km: vjust= 2.1
15km: vjust= 2.1
10km: vjust= 2.0
5km: vjust= 2.0
1km: vjust= 1.8
NOTE: also exported with more intuitive color scheme (orginal also exported with underscore original_color_)

# Same plot but transparent background
for presentations with black backgrounds

```{r}
ggplot(data=acoustic_regrid, aes(x=year, y=cm_m, color=year))+
  stat_boxplot(geom="errorbar", width=0.25, color="white")+
  geom_boxplot(data=acoustic_regrid, aes(x=year, y=cm_m), outlier.shape=NA, fill=NA, color="white")+ #plot boxplots without outliers since in jitter
  #scale_color_manual(values = c("2013"="grey20", "2014"="grey20", "2015"="chocolate", "2016"="chocolate", "2017"="grey20", "2018"="grey20"))+
   geom_jitter(width=0.15, height=0, alpha=0.3, color="#999999" )+ #add actual data points to our boxplot
  geom_text(data= summary_year, aes(x = year, y = median_CM, label = paste0("CM: ", median_CM)), size = 2.8, vjust = -0.5, color="white")+ #label median 
  geom_text(data= summary_year, aes(x = year, y = median_CM, label = paste0("n: ", no_grid_cells)), size = 2.5, vjust = 1.5, color="white")+ #n
  geom_point(data = summary_year, aes(x = year, y = quantile_cm+10, label = cld_cm, color=cld_cm), size=6, alpha=1)+ # label cld
  geom_text(data = summary_year, aes(x = year, y = quantile_cm, label = cld_cm), vjust=2.1, color="white")+ # label significance cld
  scale_color_manual(values=clrs)+ #use manual color palette defined above 
  scale_y_reverse()+
  xlab("Year")+
  ylab("CM (m)")+
  ggthemes::theme_few()+
  theme(legend.position = 'none')+
  theme(legend.position = 'none',
        panel.background = element_rect(fill = 'transparent', color="white"),
        plot.background = element_rect(fill='transparent', color="white"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.background = element_rect(fill='transparent'),
        legend.box.background=element_rect(fill = 'transparent'),
        axis.text = element_text(colour="white"),
        axis.line = element_line(color = "white"),
        axis.title = element_text(color="white"))
  
ggsave(filename=paste0("./plots/analysis 1/boxplot_kw_mc_cm_transparent", grid_size, "_.pdf"), width=7, height=5, units="in")
```


# Now re-run for each of the additional grid cell sizes
1km, 5km, 10km, 15km, 20km, 25km

# When complete, move to Ocean data.Rmd (for extracting ocean data), then satellite.Rmd (satellite light data), Ocean regrid.Rmd (regridding ocean data)






















