---
title: "Environmental variables plots"
output: html_notebook
---

# Objective of code:
In this final .Rmd following Analysis 2 GAM.Rmd, we are interested in plotting our study region, calculating summary stats about our acoustic tracks and plotting the location various ctd data 

- bathymetry map of acoustic tracks (where did we sample?)
- each oceanographic variable plotted per year

# Background
We would like to have a plot showing the larger study are for reference as well as the oceanogrpahic variables we sampled over this region

# Study area-- Map of acoustic tracks

# Add bathymetry 
Using marmap package, load NOAA bathymetry data
NOTE: these data are also loaded in Read Clean.Rmd
```{r}
library(marmap) # package for loading bathy data

# select for our region
range(echo_integration_all$Lat_M) #36 to 38.5
range(echo_integration_all$Lon_M) # -125 to -121

bathy <- getNOAA.bathy(lon1 = -125, lon2 = -121,
lat1 = 35, lat2 = 40, resolution = 1)

?getNOAA.bathy # for documentation

plot(bathy)

#convert to xyz data
bathy_xyz <- as.xyz(bathy)

#select for only ocean depths (avoid land) but while keeping V1 and V2
baty_xyz= bathy_xyz |> 
  filter(V3<=0) # land values are positive while ocean are negative so select for only negative values

##convert to dataframe for ggplot
#bathy_xyz_fortify= fortify(bathy_xyz)
```
# Mean depth
overall mean depth of sampled area
```{r}
bathy_acoustic_track |> 
  summarise(mean_depth=mean(depth))
```



# PLOT acoustic tracks with underlying bathymetry data
The following plot:
- plots bathymetry contour with specific depths displayed
- acoustic track overlay
- contour of specific 550 m selection depth (used data deeper than 550m bottom depth) *note there is one spot at the top of the map (see 2013) where tracks appear to be on right of line. Prob due to discrepancy between acousticcally detected bottom and resolution of noaa marmap plots (or start of canyon it looks like)
```{r}

ggplot(data=bathy_xyz)+
  #geom_contour(data=bathy_xyz, aes(x=V1, y=V2, z=V3), breaks = c(-200, -1000, -3000, -4000), color="#56B4E9" , linewidth=0.09)+ #Depth breaks
  annotate(geom="text", x=-122.1, y=36, label="725 m", color="#0072B2", size=3)+ #label contour
  geom_point(data= echo_integration_all, aes(x=Lon_M, y=Lat_M), shape=21, size=0.20, color=  "#D55E00" , alpha=0.65)+ #add acoustic track location info
  geom_sf(data=states, fill= "#999999" , alpha=0.25, color="black", lwd=0.25)+ #state outline
  coord_sf(xlim=c(-124.5, -121 ), ylim=c(36 ,38.5))+ #study region "core" area
  geom_contour(data=bathy_xyz, aes(x=V1, y=V2, z=V3), breaks = c(-725), color="#0072B2", size=0.5)+ # depth minimum value of 725m
  xlab("")+
  ylab("")+
  theme(axis.text.x  = element_text(angle=90),
        panel.grid = element_blank(),
        panel.background = element_rect(fill="white"))+
  facet_wrap(~year)

#ggsave("./plots/coverage.pdf", dpi=300) #pdf
```

# Basemaps
```{r}
library("sf")
library("rnaturalearth")
library("rnaturalearthdata")


coast <- rnaturalearth::ne_coastline(scale = "medium", returnclass = "sf")
countries <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
states <- rnaturalearth::ne_states(returnclass = "sf")
?ne_states()

# to view
ggplot() +
  geom_sf(data= coast)
```

# Plot: zoomed out study area
```{r}

ggplot()+
  geom_sf(data=states, fill= "#999999" , alpha=0.25, color="black", lwd=0.25)+
  geom_point(data= echo_integration_all, aes(x=Lon_M, y=Lat_M), shape=21, size=0.20, color= "#D55E00" , alpha=0.65)+ #add acoustic track location info
  annotate(geom="text", x=-120.5, y=37.75, label="San Francisco", color="black", size=3)+ # sf
  annotate(geom="point", x=-122.44, y=37.75, color="#0072B2", size=2, shape=8)+ 
  annotate(geom="text", x=-120.75, y=36.55, label="Monterey", color="black", size=3)+ 
  annotate(geom="point", x=-121.89, y=36.55, color="#0072B2", size=2, shape=8)+ 
  annotate(geom="text", x=-122.5, y=40.44, label="Cape Mendocino", color="black", size=3)+ 
  annotate(geom="point", x=-124.4, y=40.44, color="#0072B2", size=2, shape=8)+
  annotate(geom="text", x=-118.5, y=32.6, label="San Diego", color="black", size=3)+ 
  annotate(geom="point", x=-117.16, y=32.7, color="#0072B2", size=2, shape=8)+
  coord_sf(xlim=c(-125, -117 ), ylim=c(32 ,42))+ #study region "core" area
  xlab("")+
  ylab("")+
  theme(axis.text.x  = element_text(angle=90),
        panel.grid = element_blank(),
        panel.background = element_rect(fill="white"))

# save output
ggsave("./plots/analysis 2/coverage_wide.pdf",dpi=300) #pdf
```

# Start and end date of acoustic tracks per year
```{r}
echo_integration_all |> 
  group_by(year) |> 
  reframe(time_range=range(time)) #start and end dates per year
```
# CTD date range
Note that I am assuming here that the ctd_deep_data_core df is the final version of what was included in each grid cell (I believe this is the case, but its been awhile and there may have been some additional filtering). Pretty sure these are correct. 
```{r}
ctd_deep_data_core |> 
  group_by(year) |> 
  reframe(min_date= min(date_parsed, na.rm = T),
          max_date= max(date_parsed, na.rm = T)) #start and end dates per year


# Number of total ctds
acoustics_ctd |> 
  group_by(year) |> 
  summarise(no_ctds=sum(count_ctds))

acoustics_ctd |> 
  summarise(no_ctds=sum(count_ctds)) #total of 191 casts
```

# PLOT of grid cells with acoustic track and CTD position overlaid 
II update 12/1/23: getting the error message that "lon" from breaks no longer found. To remedy, 
NOTE: the change I added was in removing the grid pattern (start and end lat/lon) which was no longer available as a df. I could simply re-add these by re-running previous code from the re-gridding process, but instead I am jsut sticking with the square symbol to represent the location of grid cells (and which were matched up with lat lon in previous iterations)
```{r}

# Plot grid cells with acoustic tracks and ctd cast locations
ggplot()+
  geom_sf(data=states)+ # add a basemap
  coord_sf(xlim=c(-124.5, -121 ), ylim=c(36 ,39))+ #study region
  geom_point(data=echo_integration_all, aes(x=Lon_M, y=Lat_M), color="#D55E00" , size=0.55, alpha=0.50)+ #cells with at least 3 fine resolution cells 
  geom_point(data=ctd_deep_loc_core, aes(x=LONGITUDE, y=LATITUDE), shape=3, size=1,color="black", stroke=0.5)+ # all ctd locations
  geom_point(data=acoustics_ctd, aes(x=lon, y=lat), size=4, shape=22, color="black" , stroke=0.50, alpha=0.5)+  # grid cell 
  labs(fill="Center of mass [m]")+#change legend title
  #scale_x_continuous(breaks = lon, minor_breaks = NULL)+ #note skipping lines now because not min, max
  #scale_y_continuous(breaks=lat, minor_breaks = NULL)+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(), #tick marks from lat long coordinates
        #panel.grid = element_line(colour = "grey", linetype = "solid", linewidth =0.05),# removed gridlines but could readd here
        panel.background = element_blank())+
  labs(x="", y="")+ 
  facet_wrap(~year)

ggsave(filename = "./plots/analysis 2/coverage_grid_ctd.pdf") 
#ggsave(filename = "./plots/analysis 2/coverage_grid_ctd.png", dpi=300) #png
```
#  Plot grid cells with acoustics + ocean variables
```{r}
{ggplot()+
  geom_sf(data=states, fill="gray", color="black")+
  coord_sf(xlim=c(-124.5, -121 ), ylim=c(36 ,39))+ #study region
  geom_point(data=acoustics_ctd, aes(x=lon, y=lat, fill=cm_m), shape=22, size=3.5, alpha=0.9)+ # cells
  scale_fill_viridis_c(direction=-1)+ # scale center of mass depth
  geom_point(data=ctd_deep_loc_core, aes(x=LONGITUDE, y=LATITUDE), shape=1, size=1,color="white", stroke=0.5)+ # all ctd locations
  labs(x="", y="", fill="CM")+ 
  theme(axis.text.x  = element_text(angle=90),
        panel.grid = element_blank(),
        panel.background = element_rect(fill="white"))+
  facet_wrap(~year)}
  
ggsave(file="./plots/analysis 2/ctd_grid.pdf")
  
```

# Plot of oceanographic variable over space -- center of mass depth
The following creates a faceted map by year of center of mass depth... or other oceanographic variables for comparison

```{r}
ggplot()+
  geom_sf(data=states)+ # add a basemap
  coord_sf(xlim=c(-124.5, -121 ), ylim=c(36 ,39))+ #study region
  geom_point(data=acoustics_ctd, aes(x=lon, y=lat, fill=cm_m), size=3.8, shape=22)+  # Z is the depth at which light is 1% of surface
  cmocean::scale_fill_cmocean(name="deep", direction=1)+
  geom_text(data=acoustics_ctd, aes(x=lon, y=lat, label=round(cm_m, 0)), size=1)+  # label cm depth
  guides(fill=guide_colourbar(reverse = TRUE))+ #reverse color ramp direction in legend so deeper at bottom
  labs(fill="Center of mass [m]")+#change legend title
  theme(panel.background = element_blank(),
        axis.text.x = element_text(angle=90))+
  xlab("")+
  ylab("")+
  facet_wrap(~year)

ggsave(filename = "./plots/analysis 2/oceano/cm.pdf")
```



# Depth at 1% light 
```{r}
ggplot()+
  geom_sf(data=states)+ # add a basemap
  coord_sf(xlim=c(-124.5, -121 ), ylim=c(36 ,39))+ #study region
  geom_point(data=acoustics_ctd, aes(x=lon, y=lat, fill=Z), size=3.80, shape=22)+  # Z is the depth at which light is 1% of surface
  cmocean::scale_fill_cmocean(name="haline", direction=-1)+ #reverse color direction 
  guides(fill=guide_colourbar(reverse = TRUE))+ #reverse color ramp direction in legend so deeper at bottom
  geom_text(data=acoustics_ctd, aes(x=lon, y=lat, label=round(cm_m, 0)), size=1)+  # label cm depth
  theme(panel.background = element_blank(),
        axis.text.x = element_text(angle=90))+
  xlab("")+
  ylab("")+
  facet_wrap(~year)


# save output
ggsave("./plots/analysis 2/oceano/Z.pdf") # will auto adjust height 

```

#chlorophyll in upper 100m
```{r}
ggplot()+
  geom_sf(data=states)+ # add a basemap
  coord_sf(xlim=c(-124.5, -121 ), ylim=c(36 ,39))+ #study region
  geom_point(data=acoustics_ctd, aes(x=lon, y=lat, fill=chl_100m), size=3.8, shape=22)+  # Z is the depth at which light is 1% of surface
  cmocean::scale_fill_cmocean(name="algae", direction=1)+
  geom_text(data=acoustics_ctd, aes(x=lon, y=lat, label=round(cm_m, 0)), size=1)+  # label cm depth
  labs(fill="Chl-a")+
  guides(fill=guide_colorbar(reverse = TRUE))+
  theme(panel.background = element_blank(),
        axis.text.x = element_text(angle=90))+
  xlab("")+
  ylab("")+
  facet_wrap(~year)
    
    
    
# save output
ggsave("./plots/analysis 2/oceano/chl.pdf") # will auto adjust height 

```
#dyanmic height anomalies in 500m 

```{r}
ggplot()+
  geom_sf(data=states)+ # add a basemap
  coord_sf(xlim=c(-124.5, -121 ), ylim=c(36 ,39))+ #study region
  geom_point(data=acoustics_ctd, aes(x=lon, y=lat, fill=dyn_ht_500m), size=3.8, shape=22)+  # Z is the depth at which light is 1% of surface
  cmocean::scale_fill_cmocean(name="balance", direction=1)+
  geom_text(data=acoustics_ctd, aes(x=lon, y=lat, label=round(cm_m, 0)), size=1)+  # label cm depth
  labs(fill="Dyn ht [m]")+
  theme(panel.background = element_blank(),
        axis.text.x = element_text(angle=90))+
  xlab("")+
  ylab("")+
  facet_wrap(~year)

# save output
ggsave("./plots/analysis 2/oceano/dyn_ht.pdf") 

```
Dynamic height as measured via CTD along coast: (these are dynamic height anomalies at 500m depth)
more negative== more upwelling (greater tilt in isopycnal)
more positive== less upwelling


#oxygen at 175m
```{r}
ggplot()+
  geom_sf(data=states)+ # add a basemap
  coord_sf(xlim=c(-124.5, -121 ), ylim=c(36 ,39))+ #study region
  geom_point(data=acoustics_ctd, aes(x=lon, y=lat, fill=o2_175m), size=3.8, shape=22)+  # Z is the depth at which light is 1% of surface
  cmocean::scale_fill_cmocean(name="thermal", direction=-1)+
  geom_text(data=acoustics_ctd, aes(x=lon, y=lat, label=round(cm_m, 0)), size=1)+  # label cm depth
  labs(fill="DO at 175 m")+
  guides(fill=guide_colorbar(reverse=TRUE))+
  theme(panel.background = element_blank(),
        axis.text.x = element_text(angle=90))+
  xlab("")+
  ylab("")+
  facet_wrap(~year)

# save output
ggsave("./plots/analysis 2/oceano/o2.pdf") # will auto adjust height
```

#MVBS 
```{r}
ggplot()+
  geom_sf(data=states)+ # add a basemap
  coord_sf(xlim=c(-124.5, -121 ), ylim=c(36 ,39))+ #study region
  geom_point(data=acoustics_ctd, aes(x=lon, y=lat, fill=MVBS), size=3.8, shape=22)+  # Z is the depth at which light is 1% of surface
  cmocean::scale_fill_cmocean(name="thermal", direction=1)+
  geom_text(data=acoustics_ctd, aes(x=lon, y=lat, label=round(cm_m, 0)), size=1)+  # label cm depth
  theme(panel.background = element_blank(),
        axis.text.x = element_text(angle=90))+
  xlab("")+
  ylab("")+
  facet_wrap(~year)

# save output
ggsave("./plots/analysis 2/oceano/mvbs.pdf") 
```
# Temperature at cm depth (322 m)
```{r}
ggplot()+
  geom_sf(data=states)+ # add a basemap
  coord_sf(xlim=c(-124.5, -121 ), ylim=c(36 ,39))+ #study region
  geom_point(data=acoustics_ctd, aes(x=lon, y=lat, fill=temp_cm), size=3.8, shape=22)+  # Z is the depth at which light is 1% of surface
  cmocean::scale_fill_cmocean(name="thermal", direction=1)+
  geom_text(data=acoustics_ctd, aes(x=lon, y=lat, label=round(cm_m, 0)), size=1)+  # label cm depth
  labs(fill="Temp at 314 m [C]")+
  guides(fill=guide_colorbar(reverse = TRUE))+
  theme(panel.background = element_blank(),
        axis.text.x = element_text(angle=90))+
  xlab("")+
  ylab("")+
  facet_wrap(~year)

# save output
ggsave("./plots/analysis 2/oceano/temp_314m.pdf", dpi=300) # will auto adjust height
```
# Oceanographic variables between years

# Calculate mean values 
```{r}

# calculate median values for labeling plot
summary_yrs= acoustics_ctd |> 
               group_by(year) |> 
               summarise(median_cm=median(cm_m),
                         mean_cm=mean(cm_m),
                         median_dyn_ht=median(dyn_ht_500m), 
                         mean_dyn_ht=mean(dyn_ht_500m),
                         median_MVBS=median(MVBS),
                         mean_MVBS=mean(MVBS),
                         median_temp_10m=median(temp_10m),
                         mean_temp_10m=mean(temp_10m),
                         median_temp_175m=median(temp_175m),
                         mean_temp_175m=mean(temp_175m),
                         median_Z=median(Z),
                         mean_Z= mean(Z),
                         median_o2_175m= median(o2_175m),
                         mean_o2_175m=mean(o2_175m),
                         median_o2_500m=median(o2_500m),
                         mean_o2_500m=mean(o2_500m),
                         median_chl=median(chl_100m),
                         mean_chl=mean(chl_100m),
                         n=n()) #calculate the mean value per year

summary_yrs$year= as.factor(summary_yrs$year) # convert year to a factor
class(summary_yrs$year)
```

# Dynamic height between years
```{r}

ggplot(data=acoustics_ctd, aes(x=as.factor(year), y=dyn_ht_500m, group=year, color=as.factor(year)))+
  stat_boxplot(geom="errorbar", width=0.25)+
  geom_boxplot( aes(group=year), outlier.shape=NA, show.legend = FALSE)+ #plot boxplots without outliers since in jitter
  scale_color_manual(values = c("2013"="black" , "2014"="black", "2015"="#D55E00" , "2016"= "#D55E00" , "2017"="black", "2018"="black"))+
  geom_jitter(alpha=0.3, color="#999999" , width=0.15)+ #add actual data points to our boxplot
  geom_text(data= summary_yrs, aes(x = year, y = median_dyn_ht, label = paste0("Median:", round(median_dyn_ht, 2))), size = 2.5, vjust = -0.5)+ #label median 
  #geom_text(data= summary_yrs, aes(x =year, y = median_dyn_ht, label = paste0("n: ", n)), size = 3, vjust = 1.5, color="black")+ #label sample size 
  geom_hline(yintercept=0, linewidth=0.5, color="black", alpha=0.6)+ # 0 intercept
  xlab("Year")+
  ylab("Dynamic height anomaly [m]")+
  ggthemes::theme_few()+
  theme(legend.position = 'none')

ggsave(filename = "./plots/analysis 2/oceano/box_dyn_ht.pdf", width=170, height= 100,units="mm", dpi=300)
```
Interpretation: during the two warmest years, there seems to be an increase in dynamic height, indicative of decreased upwelling (negative values= more upwelling), 0-+ values= less upwelling, especially in 2016. 



Note: need acosutics_ctd to be numeric for other plotting so convert with caution or just switch back to numeric

# Light between years 
```{r}

ggplot(data=acoustics_ctd, aes(x=as.factor(year), y=Z, group=year, color=as.factor(year)))+
  stat_boxplot(geom="errorbar", width=0.25)+
  geom_boxplot( aes(group=year), outlier.shape=NA, show.legend = FALSE)+ #plot boxplots without outliers since in jitter
  scale_color_manual(values = c("2013"="black" , "2014"="black", "2015"="#D55E00" , "2016"= "#D55E00" , "2017"="black", "2018"="black"))+
  geom_jitter(alpha=0.3, color="#999999" , width=0.15)+
  geom_text(data= summary_yrs, aes(x = year, y = median_Z, label = paste0("Median:", round(median_Z, 2))), size = 2.5, vjust = -0.5)+ #label
  scale_y_reverse()+ #deeper at bottom of plot easier for me to grasp
  xlab("Year")+
  ylab("Euphotic depth (Z) [m]")+
  ggthemes::theme_few()+
  theme(legend.position = 'none')

ggsave(filename = "./plots/analysis 2/oceano/box_light.pdf", width=170, height= 100,units="mm", dpi=300)
```
# Temp at meso depths (175m) between years 
```{r}

ggplot(data=acoustics_ctd, aes(x=as.factor(year), y=temp_175m, group=year, color=as.factor(year)))+
  stat_boxplot(geom="errorbar", width=0.25)+
  geom_boxplot( aes(group=year), outlier.shape=NA, show.legend = FALSE)+ #plot boxplots without outliers since in jitter
  scale_color_manual(values = c("2013"="black" , "2014"="black", "2015"="#D55E00" , "2016"= "#D55E00" , "2017"="black", "2018"="black"))+
  geom_jitter(alpha=0.3, color="#999999" , width=0.15)+
  geom_text(data= summary_yrs, aes(x = factor(year), y = median_temp_175m, label = paste0("Median:", round(median_temp_175m, 2))), size = 2.5, vjust = -0.5)+ #label
  #geom_text(data= summary_yrs, aes(x =factor(year), y = median_temp_175m, label = paste0("n: ", n)), size = 2, vjust = 1.5, color="black")+ #label sample size 
  xlab("Year")+
  ylab("Temperature at 175 m [C]")+
  ggthemes::theme_few()+
  theme(legend.position = 'none')

ggsave(filename = "./plots/analysis 2/oceano/box_temp_meso.pdf", width=170, height= 100,units="mm", dpi=300)
```
Note that the changes in temperature here were pretty minimal: less than half of a degree but then again Jerome had said that any change during this time was probably a big deal at these depths


# Temp at meso depths (175m) between years 
```{r}

ggplot(data=acoustics_ctd, aes(x=as.factor(year), y=temp_10m, group=year, color=as.factor(year)))+
  stat_boxplot(geom="errorbar", width=0.25)+
  geom_boxplot( aes(group=year), outlier.shape=NA, show.legend = FALSE)+ #plot boxplots without outliers since in jitter
  scale_color_manual(values = c("2013"="black" , "2014"="black", "2015"="#D55E00" , "2016"= "#D55E00" , "2017"="black", "2018"="black"))+
  geom_jitter(alpha=0.3, color="#999999" , width=0.15)+
  geom_text(data= summary_yrs, aes(x = factor(year), y = median_temp_10m, label = paste0("Median:", round(median_temp_10m, 2))), size = 2.5, vjust = -0.5)+ #label
  #geom_text(data= summary_yrs, aes(x =factor(year), y = median_temp_175m, label = paste0("n: ", n)), size = 2, vjust = 1.5, color="black")+ #label sample size 
  xlab("Year")+
  ylab("Temperature in upper 10 m [C]")+
  ggthemes::theme_few()+
  theme(legend.position = 'none')

ggsave(filename = "./plots/analysis 2/oceano/box_temp_surface.pdf", width=170, height= 100,units="mm", dpi=300)
```




# MVBS between years
```{r}

ggplot(data=acoustics_ctd, aes(x=as.factor(year), y=MVBS, group=year, color=as.factor(year)))+
  stat_boxplot(geom="errorbar", width=0.25)+
  geom_boxplot( aes(group=year), outlier.shape=NA, show.legend = FALSE)+ #plot boxplots without outliers since in jitter
  scale_color_manual(values = c("2013"="black" , "2014"="black", "2015"="#D55E00" , "2016"= "#D55E00" , "2017"="black", "2018"="black"))+
  geom_jitter(alpha=0.3, color="#999999" , width=0.15)+
  geom_text(data= summary_yrs, aes(x = factor(year), y = median_MVBS, label = paste0("Median:", round(median_MVBS, 2))), size = 2.5, vjust = -0.5)+ #label
  xlab("Year")+
  ylab("MVBS [dB re 1 m-1]")+
  ggthemes::theme_few()+
  theme(legend.position = 'none')

ggsave(filename = "./plots/analysis 2/oceano/box_mvbs.pdf", width=170, height= 100,units="mm", dpi=300)
```
# Chl-a
```{r}

ggplot(data=acoustics_ctd, aes(x=as.factor(year), y=chl_100m, group=year, color=as.factor(year)))+
  stat_boxplot(geom="errorbar", width=0.25)+
  geom_boxplot( aes(group=year), outlier.shape=NA, show.legend = FALSE)+ #plot boxplots without outliers since in jitter
  scale_color_manual(values = c("2013"="black" , "2014"="black", "2015"="#D55E00" , "2016"= "#D55E00" , "2017"="black", "2018"="black"))+
  geom_jitter(alpha=0.3, color="#999999" , width=0.15)+
  geom_text(data= summary_yrs, aes(x = factor(year), y = median_chl, label = paste0("Median:", round(median_chl, 2))), size = 2.5, vjust = -0.5)+ #label
  xlab("Year")+
  ylab("Chl in upper 100 m [ug/L]")+
  ggthemes::theme_few()+
  theme(legend.position = 'none')

ggsave(filename = "./plots/analysis 2/oceano/box_chl.pdf", width=170, height= 100,units="mm", dpi=300)
```
If i want to add values, need to add to summary_yrs (can do easily, but haven't)

# Oxygen
```{r}

ggplot(data=acoustics_ctd, aes(x=as.factor(year), y=o2_175m, group=year, color=as.factor(year)))+
  stat_boxplot(geom="errorbar", width=0.25)+
  geom_boxplot( aes(group=year), outlier.shape=NA, show.legend = FALSE)+ #plot boxplots without outliers since in jitter
  scale_color_manual(values = c("2013"="black" , "2014"="black", "2015"="#D55E00" , "2016"= "#D55E00" , "2017"="black", "2018"="black"))+
  geom_jitter(alpha=0.3, color="#999999" , width=0.15)+
  geom_text(data= summary_yrs, aes(x = factor(year), y = median_o2_175m, label = paste0("Median:", round(median_o2_175m, 2))), size = 2.5, vjust = 1.5)+ #label
  xlab("Year")+
  ylab("Oxygen at 175 m [ml/l]")+
  ggthemes::theme_few()+
  theme(legend.position = 'none')

ggsave(filename = "./plots/analysis 2/oceano/box_DO.pdf", width=170, height= 100,units="mm", dpi=300)
```



# DESCRIPTION OF OCEANOGRAPHIC VARIABLES

I would like to write up a short paragraph in the results describing ranges and extremes of oceanographic variables from our analysis 
acoustics_ctd: df with mean values per acoustic grid cell (25km)
# Oceanographic variability
```{r}
# Light Z
range(acoustics_ctd$Z) # 7-71
# determine year of min and max values
acoustics_ctd[acoustics_ctd$Z==min(acoustics_ctd$Z),"year"] #2014
acoustics_ctd[acoustics_ctd$Z==max(acoustics_ctd$Z),"year"] #2015

# MVBS
range(acoustics_ctd$MVBS) # -75 -- -62 dB
# determine year of min and max values
acoustics_ctd[acoustics_ctd$MVBS==min(acoustics_ctd$MVBS),"year"] # 2017
acoustics_ctd[acoustics_ctd$MVBS==max(acoustics_ctd$MVBS),"year"] # 2015

# Temp at 314 m
range(acoustics_ctd$temp_175m) # 6.315 to 8.00 )314 m
# determine year of min and max values
acoustics_ctd[acoustics_ctd$temp_cm==min(acoustics_ctd$temp_cm),"year"] # 2016
acoustics_ctd[acoustics_ctd$temp_cm==max(acoustics_ctd$temp_cm),"year"] # 2014

# Dynamic height anomalies
range(acoustics_ctd$dyn_ht_500m) # -0.0867 to 0.1300
# determine year of min and max values
acoustics_ctd[acoustics_ctd$dyn_ht_500m==min(acoustics_ctd$dyn_ht_500m),"year"] # 2013
acoustics_ctd[acoustics_ctd$dyn_ht_500m==max(acoustics_ctd$dyn_ht_500m),"year"] # 2016


# DO at 150 m
range(acoustics_ctd$o2_150m) # 1.57 - 4.24
# determine year of min and max values
acoustics_ctd[acoustics_ctd$o2_150m==min(acoustics_ctd$o2_150m),"year"] # 2013
acoustics_ctd[acoustics_ctd$o2_150m==max(acoustics_ctd$o2_150m),"year"] # 2013


# Chl-a upper 100 m
range(acoustics_ctd$chl_100m) # 0.00 - 405.11
# determine year of min and max values
acoustics_ctd[acoustics_ctd$chl_100m==min(acoustics_ctd$chl_100m),"year"] # 2018
acoustics_ctd[acoustics_ctd$chl_100m==max(acoustics_ctd$chl_100m),"year"] # 2014

```



# end...



































