---
title: "R Notebook"
output: html_notebook
---
This code was written by II for the manuscript: https://doi.org/10.1093/icesjms/fsae129
This code is preceded by "satellite.Rmd" and followed by "Analysis 2 plots.Rmd"

# Objective of code:
- checks for correlation among ocean variables 
- runs a GAM model with ocean covariates
- evaluates model fit 

86 total 25 by 25 km cells that have all oceanographic data and ocean covariates (df: acoustics_ctd)


# Background:

Interested in which oceanographic variables may be influencing center of mass depth (vertical position) of mesopelagic fishes. Using a combination of CTD  and satellite (light) data, construct a generalized additive model (GAM) to determine whether the variables influenced center of mass depth.  


# gratia
install.packages("gratia",repos = c("https://gavinsimpson.r-universe.dev","https://cloud.r-project.org"))

# load packages

```{r load libraries}

library(mgcv) # GAM function
library(gratia) # gam plotting by Gavin Simpson (course instructor) - need to reload for most recent version
library(tidyverse)

```

# Look at the distribution of our response variable: CM Center of Mass (m)

```{r distribution of CM values}

acoustics_ctd |> 
  ggplot()+
  geom_density(aes(x=cm_m))+
  #geom_histogram(aes(x=CM), color="black", fill="white")+
  #facet_wrap(~Year)+
  theme_classic()
  
  # could also create a q-q plot 
qqnorm(acoustics_ctd$cm_m)
qqline(acoustics_ctd$cm_m)
```
# Pairplot
To evaluate linear correlation within our data

# pairs plot via ggplot
```{r pairs plot}

ocean_corr= acoustics_ctd |> 
  dplyr::select( MVBS, temp_175m, temp_500m, chl_100m, den_175m, den_500m, dyn_ht_500m, o2_175m, o2_500m, Z) #note that DEPTH_o2_05 is missing data points so might be causing errors

pairs(ocean_corr) # base R, hard ot read

# Install package
library(GGally)

ggpairs(ocean_corr, upper = list(continuous=wrap("cor", size=2)))+
  theme(strip.text.x = element_text(size = 5),
           strip.text.y = element_text(size = 5),
        axis.text = element_text(size=5)) #change label size

ggsave("./plots/analysis 2/paris.pdf")
```
Remove those terms with correlation coefficients >0.7 based on Dormann

This leaves:
MVBS, temp_175m, temp_500m, chl_100m, dyn_ht_500m, o2_175m, o2_500m, Z
(we removed density at 175 m and denstiy at 500 m)

# FULL MODEL-- fit a GAM -- prior to concurvity evaluation (concurvity describes non linear correlation)
Fit our full model with all of the covariates that passed our colinear examination
while noting that we will need to additionally remove covariates that approximate eachother (concurvity)
```{r}
colnames(acoustics_ctd)
class(acoustics_ctd$dyn_ht_500m) #scientific notation appears to cause problems
acoustics_ctd$dyn_ht_500m <- format(acoustics_ctd$dyn_ht_500m, scientific=FALSE)
acoustics_ctd$dyn_ht_500m <- as.numeric(acoustics_ctd$dyn_ht_500m)
acoustics_ctd$dyn_ht_500m <- round(acoustics_ctd$dyn_ht_500m, 4)

# Full GAM model:
m_full <- gam(cm_m ~ 
                s(temp_175m) + #temp at 175m
                s(temp_500m)+ #temp at 500m
                s(o2_175m) +  # oxygen at 150m
                s(o2_500m)+ #oxygen at 500m
                s(chl_100m) + # integrated chl in upper 100m
                s(dyn_ht_500m) + # dynamic height anomaly 
                s(MVBS)+ #overall backscatter/biomass 
                s(Z), #depth at which light is 1% of surface
    data = acoustics_ctd, method = "REML", family = gaussian(), select = TRUE) #select=TRUE provides significance tests for each covariate

summary.gam(m_full)
#default value is k=10, check that this is an appropriate value with k.check
```
A note on covariate selection:

Based on the pairs plot above, I removed variables that had a correlation coefficient of 0.7 or greater (although some were close to this). 
Will now check all of these correlations with concurvity feature of GAM output 

# View output of model
format(2e-16,scientific=FALSE) # to see p-values in non scientific notation
Customizing plots with gratia
https://gavinsimpson.github.io/gratia/articles/custom-plotting.html
```{r}
# MODEL fit: amount explained, significant covariates
summary(m_full)  # model summary 
```
# PLOT relationship between covariates and center of mass depth :: DRAW
For reference: mgcv::plot.gam(m_full) #plot with mgcv package

```{r}
# PLOT: relationship of covariate to cm depth
gratia::draw(m_full)& # plot with gratia
ggthemes::theme_few() #note use of & in plot to apply theme options to all plots instead of just one 

```
# Plot with {mgcv}
```{r}
plot(m_full, rug = TRUE, residuals = TRUE, se=TRUE, shade=TRUE, shade.col="cadetblue",
     pch = 1, cex = 0.5) #this includes points for residuals 
```

# Examine model fit via {gratia}
```{r}
appraise(m_full, method = "simulate") # look at plot diagnostics

# Our data appear to be approximately Gaussian, good fit except for at the extreme tails, overall a reasonable fit to our data
```

#  Further examine correlation (non-linear) amongst variables using the concurvity function on our output
first have to rule out any additional correlation among covariates

```{r}
# overall concurvity
gratia::concrvity(m_full, pairwise = FALSE) # if less than 0.8 all terms ok
# if greater than 0.8, run pairwise comparisons

conc <- concrvity(m_full, pairwise = TRUE)
conc

draw(conc)+
  geom_text(aes(label=round(.concurvity, 1)), size=2)+ #add actual values ot plot sweet none above 0.5
  theme(axis.text.x = element_text(angle=90))

ggsave("./plots/analysis 2/concurvity.pdf")
```
# Helpful references
https://gavinsimpson.github.io/gratia/reference/model_concurvity.html
https://stats.stackexchange.com/questions/580791/interpretation-of-estimate-concurvity-vs-worst-concurvity-in-gams for interpretation
NOAM ROSS: https://noamross.github.io/gams-in-r-course/chapter2

NROSS: "in general, we avoid putting multiple collinear variables into the same model" 
"with GAMS we have an additional potential pitfall. Even if two variables aren't collinear, they may have concurvity, that is, one may be a smooth curve of another" If we use both variables as predictors in a model, we get smooths with wild confidence intervals

Based on this advice, we then have to additionally remove covariates:
temperature at 500m

#More on concurvity 
```{r}
concurvity(m_full, full = TRUE) #overall - if any values high, run full=FALSE for pairwise comparisons 
concur=concurvity(m_full, full = FALSE) # pairwise concurvities
concur$worst #look at output for "worst" value 
# convert format
format(concur$worst, scientific = FALSE) #convert from scientific notation but harder to read look for values greater than 0.5 

```
NOAM ROSS: "since concurvity is complex, the function reports three different ways of measuring concurvity. Each is better in some situations. What is important is that you should always look at the worst case, and if the value is high (say, over 0.8), inspect your model more carefully"
(removed both Temperature at 500 m and Temperature at 10m)


The following is the final list of covariates: between oxygen and temperature, oxygen makes more sense because there is some evidence that biologically limiting

o2_175m
temp_175m
chl_100m
dyn_ht_500m
MVBS
Z

# FINAL MODEL -GAM
Redo the procedure from above but with set of non-correlated covariates (now all non correlated (checked both colinarity and concurvity))
 
```{r}

# modified GAM model: use this one after convo with Jerome!!! 
m_final <- mgcv::gam(cm_m ~ 
                s(temp_175m) +  # temperature at 175m depth
                s(o2_175m)+ # oxygen at 175 m
                s(o2_500m)+ # oxygen at 500 m
                s(chl_100m) + # integrated chl
                s(dyn_ht_500m) + # dynamic height anomaly 
                s(MVBS)+ #overall backscatter/biomass 
                s(Z), #depth at which light is 1% of surface note could also make this a linear term with Z instead of s(Z), same output 71.4%
    data = acoustics_ctd, method = "REML", family = gaussian(), select = TRUE)

# note all bs="tp" basis functions are thin-plate splines by default and appropriate for our analysis

summary(m_final)
```
Model explained ~81% of variability in center of mass depth (m)

```{r}
conc <- concrvity(m_final, pairwise = TRUE)
conc

draw(conc)+
  geom_text(aes(label=round(.concurvity, 1)), size=2)+ #add actual values ot plot sweet none above 0.5
  theme(axis.text.x = element_text(angle=90))
```


# Summary of final model
```{r}
summary(m_final)
mgcv::summary.gam(m_final) #same same just via {mgcv}
```

# Check model fit
```{r}
# {mgcv} model fit
mgcv::gam.check(m_final)

# console output: Full convergence after 11 iterations (check! for full convergence!)

# {gratia} model fit - same plots in ggplot format
gratia::appraise(m_final)& #gratia version of gam.check 
theme(panel.border = element_rect(color="black", fill=NA),
      panel.background = element_blank())

#save
ggsave(filename=paste0("./plots/analysis 2/m_final_fit.pdf"), width=7, height=5, units="in", dpi=300)
# NOTE: a bit of a curve in lower left that falls outside of confidence interval. Also a couple points in the extremes outside conf int but overall, I am goign to go with prett decent model fit! 
```
NOTE: use & instead of + to add theme elements to plots in {gratia} (otherwise it will only apply changes to last plot)

# check on k (number of basis functions or wiggliness) that go into smooth function
```{r}

# {mgcv}
mgcv::k.check(m_final) #appropriate k # of smooths

# all p-values non-significant, so appropriate k values (# basis functions) for our data
```
# Check (double check): concurvity with modified version of full model
All below recommended values
```{r}
# Test for concurvity
# ensure "worst" concurvity for all terms < ~0.8
conc_full=gratia::concrvity(m_final, pairwise = FALSE)
conc_full
format(conc_full$concurvity, scientific=FALSE)>0.8 # all below 0.8 (but just barely), run pairwise and examine as well


# Pairwise concurvity
conc_final <- concrvity(m_final, pairwise = TRUE)

# plot output of 
draw(conc_final)+
  geom_text(aes(label=round(concurvity, 2)), size=2)+ #add actual values ot plot sweet none above 0.5
  theme(axis.text.x = element_text(angle=90))
```

# Partial effects plots; the effect of a given covariate on center of mass depth
ALL oceanographic covariates 

```{r}
#plot partial effects plots
gratia::draw(m_final)&
  ggthemes::theme_few()

ggsave("./plots/analysis 2/gam_partialeffects_all.pdf", width=10, height=8, dpi=300)
```
# Select for significant covariates

```{r}
#select our significant covariates
p1 <- draw(m_final, select = "s(dyn_ht_500m)")
p2 <- draw(m_final, select = "s(MVBS)")
p3 <- draw(m_final, select = "s(Z)")

# Plot side by side
p3+p2+p1+patchwork::plot_layout(ncol=4) & ggthemes::theme_few()

# Update individual plot: dynamic ht anomalies
p1=p1+
  ggtitle("")+
  xlab("Dynamic ht anomaly [m]")+
  ylab("")+
  scale_x_continuous(expand = c(0,0))+
  ggthemes::theme_few()+
  theme(axis.title.x = element_text(size=16))

p1

ggsave("./plots/analysis 2/partial_effects_dyn_ht.pdf", width=85, height=75, units="mm", dpi=300)

# Update individual plot: MVBS
p2=p2+
  ggtitle("")+
  xlab("MVBS [dB re 1 m-1]")+
  ylab("")+
  scale_x_continuous(expand = c(0,0))+
  ggthemes::theme_few()+
  theme(axis.title.x = element_text(size=16))

ggsave("./plots/analysis 2/partial_effects_mvbs.pdf", width=85, height=75, units="mm", dpi=300)


# Update individual plot: Euphotic depth
p3=p3+
  ggtitle("")+
  xlab("Euphotic depth (Z) [m]")+
  ylab("")+
  #ylab("Partial effect on center of mass depth")+
  scale_x_continuous(expand = c(0,0))+
  ggthemes::theme_few()+
  theme(axis.title.x = element_text(size=16))
ggsave("./plots/analysis 2/partial_effects_z.pdf", width=85, height=75, units="mm", dpi=300)

# Plot side by side
p3+p1+p2+patchwork::plot_layout(ncol=3) & ggthemes::theme_few()

# save output
ggsave("./plots/analysis 2/partial_effects_significant.pdf", width=170, height=100, units="mm", dpi=300)
```
 "the smooth effect of the covariate on the response conditional upon the other estimated terms" - G. Simpson
 The zero line means the overall mean (on the link scale) of the response; negative values on the axis indicate where the effect of the covariate reduces the response below the average value, and positive values on the axis indicate those covariate values where the response is increased above the average 

 
# PLOT partial effects plots with mgcv instead of gratia
mgcv::plot.gam
https://rdrr.io/cran/mgcv/man/plot.gam.html
select: select an individual plot (ordered 1-# of covariates)
scale: -1 to have the same y-axis scale for each plot
ylab="Parital effect"
shade: 2 se confidence bands
seWithMean:	
if TRUE the component smooths are shown with confidence intervals that include the uncertainty about the overall mean. If FALSE then the uncertainty relates purely to the centred smooth itself. If seWithMean=2 then the intervals include the uncertainty in the mean of the fixed effects (but not in the mean of any uncensored smooths or random effects). Marra and Wood (2012) suggests that TRUE results in better coverage performance, and this is also suggested by simulation.

```{r}

# plot all
#plot.gam(m_final) #defaults

# save plot
pdf(file = "./plots/analysis 2/partial_effects.pdf", width = 14, height=5, bg="white")

# Plot individual plots -- Temperature
{par(mfrow=c(1,4), family="Helvetica",  mai=c(1, 0.55, 0.25, 0.25),oma=c(0, 2, 0,0))
# dynamic height
plot.gam(m_final, rug = TRUE, residuals = FALSE, se=TRUE, shade=TRUE, shade.col="grey",
      scale=-1,  ylab="", seWithMean=TRUE, select=4, xlab = "Dynamic Height [m]",cex.axis=1.75, cex.lab=2)
title(ylab="Partial Effect", line=2.5 , cex.lab=2)

# Light
plot.gam(m_final, rug = TRUE, residuals = FALSE, se=TRUE, shade=TRUE, shade.col="grey",
     scale=-1,  ylab="", seWithMean=TRUE, select=6, xlab = "Z (Depth Light) [m]",cex.axis=1.75, cex.lab=2)
# MVBS
plot.gam(m_final, rug = TRUE, residuals = FALSE, se=TRUE, shade=TRUE, shade.col="grey",
      scale=-1,  ylab="", seWithMean=TRUE, select=5, xlab = "Mean Vol Backscatter [dB]",cex.axis=1.75, cex.lab=2)
 #temp at cm
plot.gam(m_final, rug = TRUE, residuals = FALSE, se=TRUE, shade=TRUE, shade.col="grey",
      scale=-1, seWithMean=TRUE, select=2, ylab="", xlab = "Temp at 314 m [C]", cex.axis=1.75, cex.lab=2)
}

dev.off()
```



# NEXT...
Final Rmd. Analysis 2 plots.Rmd





















