---
title: "Loading cleaning CTD data from RREA"
output: html_notebook
---

# Objective of code
This project was created to load CTD data from the RREAS survey and to select those data corresponding to our acoustic tracks. These data will then be used to determine which environmental drivers might be responsible for affecting the acoustic Center of Mass (M) depth.

- load CTD data (metadata (ctd_header) + vertical profile data (ctd_cast))- and clean and link these tables
- select for CTD stations that recorded data to 500m
- select for physical variables of interest at specific depths of interest (150m, CM depth 314 m, 500m)


# Load libraries:
```{r, echo=FALSE}
library(tidyverse) # data manipulation, plotting
library(lubridate) #dates
library(sf)# mapping
library(oce) # visualizing and calculating values from profiles
library(ggthemes)
```

# Plot acoustic tracks
We can create a simple plot of the location of our acoustic backscatter data from the echo_integration data Lat and Lon data which represents the center point of each 100m horizontal water column (from 150m-525m) of data. Note that this is prior to re-gridding into 25 by 25km grid cells 

```{r plot acoustic tracks}

ggplot()+
  geom_sf(data=states, fill="grey", color="black")+# plot state background
  geom_point(data = echo_integration, aes(x=Lon_M, y=Lat_M, color=MVBS), size=2, shape= 21)+ #Plot location of cruisetracks
  cmocean::scale_color_cmocean(name="thermal")+ #color scale
  coord_sf(xlim = c(min(echo_integration$Lon_M-0.25), max(echo_integration$Lon_M+0.25)), ylim = c(min(echo_integration$Lat_M-0.25), max(echo_integration$Lat_M+0.25)))+ # adjust coord to min and max coord this is our "zoom"
  theme(panel.background = element_blank(), 
        #plot.background = element_blank(), #removes background of plot area
        legend.key = element_blank(),
        axis.text.x = element_text(angle=90))+ #remove grey around symbols in legend
  labs(x="Longitude", y="Latitude", fill = "Sv from RREA survey")+ #legend
  guides(fill = guide_colorbar(reverse=T))+ # reverse continuous variable legend so deeper at bottom
  facet_wrap(~year)

ggsave("./plots/mvbs_acoustic_track.pdf")
```

```{r plot acoustic tracks CM}

ggplot()+
  geom_sf(data=states, fill="grey", color="black")+# plot state background
  geom_point(data = echo_integration, aes(x=Lon_M, y=Lat_M, color=CM), size=2, shape= 21)+ #Plot location of cruisetracks
  cmocean::scale_color_cmocean(name="deep")+ #color scale
  coord_sf(xlim = c(min(echo_integration$Lon_M-0.25), max(echo_integration$Lon_M+0.25)), ylim = c(min(echo_integration$Lat_M-0.25), max(echo_integration$Lat_M+0.25)))+ # adjust coord to min and max coord this is our "zoom"
  theme(panel.background = element_blank(), 
        #plot.background = element_blank(), #removes background of plot area
        legend.key = element_blank(),
        axis.text.x = element_text(angle=90))+ #remove grey around symbols in legend
  labs(x="Longitude", y="Latitude", fill = "Sv from RREA survey")+ #legend
  guides(fill = guide_colorbar(reverse=T))+ # reverse continuous variable legend so deeper at bottom
  facet_wrap(~year)

ggsave("./plots/cm_acoustic_track.pdf")
```

# Load CTD data from RREA survey into R  
These data are currently stored in the folder ctd_data
```{r load CTD files}

# Read ctd header info (ctd metadata)
ctd_header <- read.csv(file="./ctd_data/dbo_CTD_HEADER.txt", header = TRUE, sep=",", stringsAsFactors = FALSE)
#add stringsAsFactors=FALSE to solve later error in utf-8

# load CTD cast info (vertical oceanographic data)
ctd_cast <- read.csv(file="./ctd_data/dbo_CTD_CAST.txt", header = TRUE, sep=",", stringsAsFactors = FALSE)

```
# Clean up imported CTD data- select for survey years of interest (2013:2018)
-convert lat lon
-col for year
*Note CTD_LAT and CTD_LON are the position info for the actual locations of the CTD (not the station position that was the likely target location)
```{r clean header df}

# clean up CTD data
ctd_header <- ctd_header %>% 
            mutate(date_parsed= mdy_hms(CTD_DATE)) %>% # create a new variable (column) parsing the date from factor to POSIX ct and t
            mutate(Year= year(date_parsed)) %>% # create a new column "year" which extracts the year info from our parsed date column
            mutate(day_ctd= as_date(date_parsed))%>% #create a new col with day
            rename(y = CTD_LAT, x = CTD_LONG) %>% # rename our latitude (y) and longitude (x) columns
            mutate(lat.d  = as.numeric(str_sub(y, start=1, end= 2)), #create a new column of first 2 terms dd from latitude
            lat.m  = as.numeric(str_sub(y, start=3, end=6)), # extract minutes terms (mm.mm) from 3-6 position in lat col
            LATITUDE    = lat.d + lat.m/60, # divide mm.mm by 60 and add to dd terms
            long.d = as.numeric(str_sub(x, start=1, end= 3)), #first 3 terms ddd
            long.m = as.numeric(str_sub(x, start=4, end=7)), #extract mm.mmm terms in position 4-7 
            LONGITUDE = -(long.d + long.m/60)) %>% # create new col for longitude, note - for longitude values
            dplyr::select(-c(lat.d, long.d, lat.m, long.m, x, y)) |>  #remove intermediary columns 
            rename(year=Year)

#create vector of years of our study
survey_yrs <- 2013:2018

# Select for our years of interest: 2013-2018
 ctd_header <- ctd_header %>%
                filter(year %in% survey_yrs)

#the code above converts CTD_LAT to LATITUDE and CTD_LONG to LONGITUDE variables in decimal degrees- this is the specific position of a given CTD, not the station position. Note there is a value for each record.
 
 # select for our years of interest (2013-2018) for the vertical profile info:Because we haven't merged these data to date cols, can only select by CRUISE ID. Note this is just in an effort to reduce the size of the df (a lot of data with all years + all casts)
 
 
 cruise_yrs <- c(1305, 1405, 1505, 1603, 1703, 1802)
 
ctd_cast<-  ctd_cast %>%
   filter(CRUISE %in% cruise_yrs)
 
# this selects for the vertical profile data for the years 2013--2018
```

# Plot acoustic tracks + CTD locations (all CTDs, regardless of depth)
```{r plot acoustic tracks with CTD locations}
ggplot()+
  geom_sf(data=states)+
  geom_point(data=echo_integration, aes(x=Lon_M, y=Lat_M, color=year), show.legend = F)+ #plot acoustic tracks
  geom_point(data=ctd_header, aes(x= LONGITUDE, y=LATITUDE), color="black", shape=3,  size=1, show.legend = F)+
  coord_sf(crs=st_crs(states), xlim = c(min(echo_integration$Lon_M-0.25), max(echo_integration$Lon_M+0.25)), ylim = c(min(echo_integration$Lat_M-0.25), max(echo_integration$Lat_M+0.25)))+
  facet_wrap(~year)+
  theme_classic()
```
# Join the actual vertical CTD data (ctd_cast) to our metadata (ctd_header for 2013:2018)

The following adds all CTD oceanographic variables to our header data for ALL (regardelss of depth) 2013:2018 sites 

```{r join CTD cast data to CTD header data}

#left join ctd_cast (CTD vertical data) to CTD station metadata (ctd_header) - in this case all CTD casts (regardless of depth, from 2013-2018)
ctd_data <- left_join(x=ctd_header, y=ctd_cast, by=c("CRUISE","CTD_INDEX")) 

# add ctd_cast to ctd_header df by the cols CRUISE and CTD_INDEX (unique casts)

# each obs (row) represents a CTD data point (for multiple variables- columns) at a particular location
rm(cruise_yrs, survey_yrs) 

```

# Determine the deepest depth a given CTD cast actually recorded vertical data to! 
Select casts that went at least 500m deep

```{r select for CTD based on deepest record}
# Create a df of those CTD casts (unique CRUISE and CTD_INDEX) with a max depth (last row per CTD_INDEX) of at last 150m

ctd_deep <- ctd_data %>% 
  group_by(CRUISE, CTD_INDEX) %>% #unique code for each year + CTD cast
  slice(n())%>% # select the last row for each unique CTD (CTD_INDEX) (and year)
  dplyr::select(CRUISE, CTD_INDEX, CTD_BOTTOM_DEPTH, CTD_DEPTH)%>%
  filter(CTD_DEPTH>= 500)

# keep only CRUISE and CTD_INDEX (going to join by these values and don't want redundant depth columns -- need to add ALL records per cast)
ctd_deep <- dplyr::select(ctd_deep, c(CRUISE, CTD_INDEX))

# select for CRUISE and CTD_INDEX and join to that df
ctd_deep_data <- left_join(x= ctd_deep, y=ctd_data, by=c("CRUISE", "CTD_INDEX")) 


# remove intervening dfs
rm(ctd_deep, ctd_data)

# remove initial tables as ctd_deep_data now has all info from 2013:2018 for those CTD casts where the recorded depth was at least 150m
rm(ctd_cast, ctd_header)

```

# Updated map of acoustic track and CTD casts -- create unique location df

```{r plot of CTD data with acoustic track one point per cast }

# create df of unique locations - since already have code for selecting the last entry per cast, going to do the same here
# this has the added benefit of having the max depth an individual CTD sampled
ctd_deep_loc <- ctd_deep_data %>% 
                group_by(CRUISE, CTD_INDEX) %>% #unique code for each year + CTD cast
                slice(n())%>% # select the last row for each unique CTD (CTD_INDEX) (and year)
                dplyr::select(CRUISE, CTD_INDEX, CTD_BOTTOM_DEPTH, CTD_DEPTH, LATITUDE, LONGITUDE, year)
```
# Plot CTD locations (casts that recorded data to at least 500m) over acoustic tracks
Plotted the ctd index value so we can remove those outside study region

```{r plot CTDs over acoustic tracks}

# Plot these locations with our acoustic track
ggplot()+
  geom_sf(data=states)+
  geom_point(data=echo_integration, aes(x=Lon_M, y=Lat_M), color="snow3", alpha=0.5, show.legend = F)+ #plot acoustic tracks
  scale_color_viridis_d()+
  geom_point(data=ctd_deep_loc, aes(x= LONGITUDE, y=LATITUDE, fill=CTD_DEPTH), color="transparent", shape=21,  size=1, show.legend = T)+
  geom_text(data=ctd_deep_loc, aes(x=LONGITUDE, y=LATITUDE, label=CTD_INDEX), size=1.5, hjust=0.5, vjust=0.5)+
  scale_fill_continuous(trans = 'reverse')+ #reverse color scale 
  coord_sf(crs=st_crs(states), xlim = c(min(echo_integration$Lon_M-0.25), max(echo_integration$Lon_M+0.25)), ylim = c(min(echo_integration$Lat_M-0.25), max(echo_integration$Lat_M+0.25)))+
  facet_wrap(~year)+
  theme_classic()

```
# Remove stations that were outside core region or beyond our range (too far offshore)
- In 2016 there were 3 stations outside of our typical sampling range, which need to be removed. (Likewise I didn't analyze acoustics data from this far out in this year). 
- in 2018 there are some additional outside regions to be removed 
-Need to also remove these stations from our vertical data: ctd_deep_data
- Would also like to restrict our geographic range of CTD samples to just be from the core region (defined as 36.45 to 38.33)
```{r}
# Remove stations from 2016 (195, 194, 193) and 2018 (182, 172, 162, 221) that were sampled far outside of our typical core region (didn't process acoustics out there either)
ctd_deep_loc_core <- ctd_deep_loc %>%
                filter(!(year==2016 & CTD_INDEX %in% c(195, 194, 193, 192)))%>% #remove stations from 2016 outside our acoustic tracks
                filter(!(year==2018 & CTD_INDEX %in% c(220, 219, 218, 217, 216, 221)))%>% # remove stations from 2018 outside acoustic tracks
                filter(LATITUDE >= 36.45 & LATITUDE <= 38.33) #core region


# Remove the same stations from the data df
ctd_deep_data_core <- ctd_deep_data %>%
                filter(!(year==2016 & CTD_INDEX %in% c(195, 194, 193, 192))) %>%
                filter(!(year==2018 & CTD_INDEX %in% c(220, 219, 218, 217, 216, 221)))%>% # remove stations from 2018 outside acoustic tracks
                filter(LATITUDE >= 36.45 & LATITUDE <= 38.33) # Select for only those CTD stations within the CORE region- LAT 36.45 to 38.33

# rm previous (more inclusive version of CTD data)
rm(ctd_deep_data, ctd_deep_loc)

```

# Plot final CTD stations with acoustic track 
The following dfs:
ctd_deep_data_core:: Vertical data for those CTD casts deployed within the CORE region and with a max depth of at least 500m (or greater) and with the couple offshore stations removed.
ctd_loc_core:: Same as above but just the last row of each CTD deployment from ctd_deep_data_core (so max depth the ctd actually visited and a single lat long so easier to plot than the vertical data)

```{r final CTDs and acoustic tracks}

# Plot these locations with our acoustic track
ggplot()+
  geom_sf(data=states, fill="bisque4", alpha=0.7)+
  geom_point(data=echo_integration, aes(x=Lon_M, y=Lat_M), color="grey", alpha=0.75, size=1, show.legend = F)+ #plot acoustic tracks
  geom_point(data=ctd_deep_loc_core, aes(x= LONGITUDE, y=LATITUDE, fill=CTD_DEPTH), color="darkblue", shape=3,  size=1.25, show.legend = F)+
  coord_sf(crs=st_crs(states), xlim = c(min(echo_integration$Lon_M-0.25), max(echo_integration$Lon_M+0.25)), ylim = c(min(echo_integration$Lat_M-0.25), max(echo_integration$Lat_M+0.25)))+
  facet_wrap(~year)+ # year not Year
  xlab("")+
  ylab("")+
  theme_classic()

ggsave(filename = "./plots/ctd/ctd_deeper_500.jpeg")
```
# 25 km gridded data + acoustic track + CTD
#E69F00
#CC79A7
```{r}
ggplot()+
  geom_sf(data=states, fill="grey", color="black")+# plot state basemap
  geom_point(data = acoustic_regrid, aes(x= lon, y=lat, fill=cm_m), color="black", stroke=0.25, size=6, shape=22, alpha=1)+ #plot regridded data
  scale_fill_viridis_c(option = "D", direction=-1)+ #color output via viridis
  geom_point(data=echo_integration_all, aes(x=Lon_M, y=Lat_M), size=0.05, color="#E69F00", alpha=0.9)+ #overlay of cruise track over regridded data
  geom_point(data=ctd_deep_loc_core, aes(x= LONGITUDE, y=LATITUDE), color="white", shape=3,  size=1.25, stroke=0.75, show.legend = F)+
  coord_sf(xlim = c(min(acoustic_regrid$lon-0.25), max(acoustic_regrid$lon+0.25)), ylim = c(min(acoustic_regrid$lat-0.25), max(acoustic_regrid$lat+0.25)))+ 
  scale_x_continuous(breaks = lon, minor_breaks = NULL)+
  scale_y_continuous(breaks=lat, minor_breaks = NULL)+
  theme(axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_blank(),
        #panel.grid = element_line(colour = "black", linetype = "solid", linewidth =0.05),
        panel.background = element_blank())+
  labs(x="", y="", fill = "CM")+ #legend
  facet_wrap(~year)
  
  
#Save plot
ggsave(filename=paste0("./plots/regrid_coverage_", grid_size, "_map.pdf"), width=170, units="mm")
```

# Plot position of CTDs with regridded acoustics data
I regridded all of the acoustics data to a 25 by 25km grid. 
Would like to visually look at these cells with CTD overlay
```{r}
ggplot()+
  geom_sf(data=states)+
  geom_point(data=acoustics_regrid_25km, aes(x=lon, y=lat), size=5.2, shape=22, alpha=0.5, color="#0072B2")+
  scale_fill_viridis_b()+
  geom_point(data=ctd_deep_loc_core, aes(x= LONGITUDE, y=LATITUDE), color="black", shape=3,  size=1.25, show.legend = F)+
  coord_sf(crs=st_crs(states), xlim = c(min(ctd_deep_loc_core$LONGITUDE-0.25), max(ctd_deep_loc_core$LONGITUDE+0.25)), ylim = c(min(ctd_deep_loc_core$LATITUDE-0.25), max(ctd_deep_loc_core$LATITUDE+0.25)))+
  scale_x_continuous(breaks = round(acoustics_regrid_25km$lon_grid_low, 2), minor_breaks = NULL)+
  scale_y_continuous(breaks=round(acoustics_regrid_25km$lat_grid_hi,2), minor_breaks = NULL)+
  facet_wrap(~year)+
  xlab("")+
  ylab("")+
  theme_classic()+
  theme(panel.grid = element_line(color="grey"),
        axis.text.x = element_text(angle=90))

ggsave(filename = "./plots/ctd/ctd_deeper_500_grid.pdf", width=8.5, dpi=300)
```
# Plot vertical profile of CTD data (ex. one indivdiual cast)

```{r Plot a practice CTD profile}
# individual CTD cast or example

ggplot(data = filter(ctd_deep_data_core, year=="2014" & CTD_INDEX==8))+ #MANUALLY SELECTED
  geom_point(aes(y=CTD_DEPTH, x= TEMPERATURE, color= "Temp"), size=1, alpha=0.4)+
  geom_line(aes(y=CTD_DEPTH, x= TEMPERATURE, color= "Temp"), linewidth=0.5, orientation = "y")+
  geom_point(aes(y=CTD_DEPTH, x= OXYGEN, color= "Oxygen"), size=1, alpha=0.4)+
  geom_line(aes(y=CTD_DEPTH, x= OXYGEN, color= "Oxygen"), linewidth=0.5, orientation = "y")+
  #geom_point(aes(y=CTD_DEPTH, x= SALINITY, color= "Salinity"), size=1)+
  #geom_line(aes(y=CTD_DEPTH, x= SALINITY, color= "Salinity"), size=1, orientation = "y")+
  scale_y_reverse()+
  scale_color_manual(values = c(
    "Temp"="firebrick4",
    "Oxygen"= "skyblue4",
    "Salinity"= "orange3"
  ))+
  labs(color="Oceanographic variable")+
  ylab("Depth (m)")+
  xlab("")+
  theme(panel.background = element_blank(),
        panel.grid = element_blank(),
        legend.key = element_blank())



# or ALL profiles per year (by CTD_INDEX)
ggplot(data = ctd_deep_data_core)+
  #geom_point(aes(y=CTD_DEPTH, x= TEMPERATURE, color= "Temp"), size=1, alpha=0.4)+
  geom_line(aes(y=CTD_DEPTH, x= TEMPERATURE, color= "Temp"), size=0.5, orientation = "y")+
  #geom_point(aes(y=CTD_DEPTH, x= OXYGEN, color= "Oxygen"), size=1, alpha=0.4)+
  geom_line(aes(y=CTD_DEPTH, x= OXYGEN, color= "Oxygen"), size=0.5, orientation = "y")+
  #geom_point(aes(y=CTD_DEPTH, x= SALINITY, color= "Salinity"), size=1)+
  #geom_line(aes(y=CTD_DEPTH, x= SALINITY, color= "Salinity"), size=1, orientation = "y")+
  scale_y_reverse()+
  scale_color_manual(values = c(
    "Temp"="firebrick4",
    "Oxygen"= "skyblue4"
  ))+
  labs(color="Oceanographic variable")+
  ylab("Depth (m)")+
  xlab("")+
  facet_wrap(~year)+
  theme(panel.background = element_blank(),
        panel.grid = element_blank(),
        legend.key = element_blank())
  
  ggsave("./ctd/allprofiles.jpg", bg="white")
```
# QAQC: Look at TS plots and see which casts have errant values
Plot TS plots for all of our years of data
```{r}

# or ALL profiles per year (by CTD_INDEX)
ggplot(data = filter(ctd_deep_data_core, year==2015), aes(x=SALINITY, y=TEMPERATURE, color=CTD_INDEX))+
  #geom_point(aes(y=TEMPERATURE, x= SALINITY, color= CTD_INDEX), size=0.5)+
  geom_text(aes(x=SALINITY, y=TEMPERATURE, label=CTD_INDEX), size=2)+
  ylab("Temperature (C)")+
  xlab("Salinity")+
  #facet_wrap(~Year)+
  theme(panel.background = element_blank(),
        panel.grid = element_blank(),
        legend.key = element_blank())


```

# FOR LOOP: Create a profile plot of each oceanographic variable per CTD cast
Export each plot to a folder of oceanographic data
Then, I went through and visually examined all of the variables (each ctd cast)
```{r}
# would like to iterate over each ctd cast per given year

yr= unique(ctd_deep_data_core$year) #vector of years
#for each CTD cast within a given year, plot a vertial profile and export

# select for a specific year (2013-2018)
for (i in 1:length(yr)){

    # filter all CTD output for year i  
    yr_ctd <- ctd_deep_data_core %>%
              filter(year %in% yr[i])
  
  
  #create vector of unique ctd casts (CTD_INDEX) for given year
    yr_ctd_cast= unique(yr_ctd$CTD_INDEX)
  
  #then for given year, iterate over each ctd cast
  for (j in 1:length(yr_ctd_cast)){
    
    
  
  #select indivdiual ctd cast within a given year
    ctd_cast_data <- yr_ctd %>% #from data for all casts in year i
                     filter(CTD_INDEX %in% yr_ctd_cast[j])#select individual ctd cast

    
    # unique oceanographic variables
    ocean_variable <- c("TEMPERATURE", "OXYGEN", "SALINITY", "CHLOROPHYLL", "DENSITY", "DYN_HGT", "IRRAD", "TRANSMISSIVITY")
    
    
    # select for each oceanogrpahic variable (col) within the specific cast and plot
    for (k in 1:length(ocean_variable)){
    
  # create a plot for specific year and ctd cast within that year-- oceanographic variable

    
  ggplot()+
    geom_line(data=ctd_cast_data, aes(y=CTD_DEPTH, x= !!sym(ocean_variable[k])), color="grey", linewidth=1.5, orientation="y")+ #line
    geom_point(data=ctd_cast_data, aes(y=CTD_DEPTH, x= !!sym(ocean_variable[k])), color="sienna3", size=0.5, alpha=0.4)+ #plot individual points above
    scale_y_reverse(expand=c(0,0))+ #remove extra space so plots begin at the surface
    xlab(paste0("Variable ", ocean_variable[k]))+
    ylab("Depth (m)")+
    #xlim(0, max(ctd_cast_data$ocean_variable[k]))+ #x axis limits based on max value of variable
    ggtitle(paste0(ocean_variable[k], yr[i], " CTD Index: ", yr_ctd_cast[j]))+ # create a title for CTD station year and CTD cast
  geom_hline(yintercept = 150, linetype="dashed", color="darkgrey")+ #simple line demarking mesopelagic depths
  geom_hline(yintercept = 500, linetype="dashed", color="darkgrey")+
  theme(panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank(), 
        panel.background = element_blank(),
        panel.border = element_rect(colour="black", fill=NA, linewidth =2),
        legend.key = element_rect(fill=NA))
# quick note on aes x input: !! removes quotes and sym() converts to symbol so can be read as raw column name
  
  # export plot with unique name
  ggsave(filename=paste0("./plots/ctd/profiles/", ocean_variable[k],"/",ocean_variable[k], "_",yr[i], "_",yr_ctd_cast[j], ".jpeg")) #store output to folder plot/ctd/ temp
    }
  }
    
}
rm(ctd_yr, ocean_variable, yr, yr_ctd_cast, i, j, k, ctd_cast_data)#remove iterative objects when all done

```

# REMOVE bad CTD data

Based on a visual inspection of all CTD profiles, I removed the following salinity and density data:
2015: cast 220 removed below 300m
2017: cast 28 removed below 100m

In both cases I removed salinity and density data (but left other variables) below these depths

```{r}

# CONVERT VALUES BELOW SPECIFIC DEPTH TO NA: 2015, CTD cast 220 remove below 300m-- salinity and density
ctd_deep_data_core$SALINITY[ctd_deep_data_core$year==2015 & ctd_deep_data_core$CTD_INDEX==220 & ctd_deep_data_core$CTD_DEPTH>=300] <- NA
ctd_deep_data_core$DENSITY[ctd_deep_data_core$year==2015 & ctd_deep_data_core$CTD_INDEX==220 & ctd_deep_data_core$CTD_DEPTH>=300] <- NA

# 2017, CTD cast 28 remove below 100m-- salinity and density
ctd_deep_data_core$SALINITY[ctd_deep_data_core$year==2017 & ctd_deep_data_core$CTD_INDEX==28 & ctd_deep_data_core$CTD_DEPTH>=100] <- NA
ctd_deep_data_core$DENSITY[ctd_deep_data_core$year==2017 & ctd_deep_data_core$CTD_INDEX==28 & ctd_deep_data_core$CTD_DEPTH>=100] <- NA

# Now we have removed the couple of outliers in our ctd data
```
 df: ctd_deep_data_core is our cleaned, deep, core CTD data (ie what we will use for all subsequent analysis)

# ADDITIONAL CTD METRICS--- CALCULATED METRICS

# CHLOROPHYLL-A : smoothing
REFERENCES:https://www.r-bloggers.com/2014/01/smoothing-ctd-profiles/

Calculated a moving average https://stackoverflow.com/questions/743812/calculating-moving-average stats::filter
```{r}
ungroup(ctd_deep_data_core) 

chl_smooth_fx= function(x, n = 3){stats::filter(x, rep(1 / n, n), sides = 2)} #function to calculate moving average of n cells


# add new column for chl smooth (which is the smoothed value of chl with cell above and below)
ctd_deep_data_core <- ctd_deep_data_core |> # all of our ctd data
                      group_by(CTD_INDEX, year) |> #so for each ctd cast per year
                      mutate(chl_smooth=round(chl_smooth_fx(x=CHLOROPHYLL, n=3), 2)) #calculate a smoothed value for chl based on 3 cells (1 on each side)

```

Note: because we used a cell size of 3 (so one neighboring value on each side), we no longer have a surface value (because no data on one side). As i am planning to simply filter for data <100m and take the depth integrated sum, this shouldn't be a problem as it was applied to each cast
(ie sum of smoothed chl-a values from 0-100m)

## Add additional vertical metrics (potential temp, potential density, spice)

# Load {oce} package:
To convert our df object to a ctd object that {oce} recognizes:
- first create a column for pressure (a requirement to be recognized)

swPressure(Depth, latitude= , eos=getOption("oceEOS", default="gsw"))
```{r oce package create ctd object}

#currently our ctd data is stored as df: ctd_deep_data_core
head(ctd_deep_data_core)

# To convert our df to a ctd object in {oce} requires a pressure value
# create column for pressure (calculated with depth and latitude):
ctd_deep_data_core$PRESSURE <- swPressure(depth=ctd_deep_data_core$CTD_DEPTH, latitude = ctd_deep_data_core$LATITUDE, eos ="gsw")
# calculation can be gsw - default but only if you have latitude info, otherwise "unesco"

# Create a ctd object in {oce}-- ALL ctd casts
ctd<- as.ctd(ctd_deep_data_core$SALINITY, ctd_deep_data_core$TEMPERATURE, ctd_deep_data_core$PRESSURE, longitude=ctd_deep_data_core$LONGITUDE, latitude= ctd_deep_data_core$LATITUDE)

```


# Calculate SPICE via {oce} package

Compute seawater "spice", also called "spiciness" (a variable orthogonal to density in TS space), in either of two formulations, depending on the value of the eos argument. If eos="unesco" then Flament's (reference 1) formulation is used. If eos="gsw" then the Gibbs SeaWater formulation for "spiciness0" is used (see reference 2).

https://rdrr.io/cran/oce/man/swSpice.html
https://doi.org/10.1139/cjfas-2017-0480
calculated spice at the 26.0 isopycnal 

Steps to calculating spice:
- Add col for potential temp
- Add col for potential density
- select for 26.0 potential density
- calculate spice (both Flament's ("unesco"), or "gsw"= Gibbs SeaWater formulation)

```{r calculate spice}
#Add col for potential temperature
ctd_deep_data_core$Potential_Temperature <- oce::swTheta(salinity = ctd_deep_data_core$SALINITY, temperature = ctd_deep_data_core$TEMPERATURE, pressure = ctd_deep_data_core$PRESSURE, referencePressure = 0, longitude = ctd_deep_data_core$LONGITUDE, latitude = ctd_deep_data_core$LATITUDE, eos = "gsw")
  
# Add col for potential density -- in situ temp (auto converted)-- note used Potential Temperature
ctd_deep_data_core$Potential_Density <- oce::swRho(salinity = ctd_deep_data_core$SALINITY, temperature = ctd_deep_data_core$Potential_Temperature, pressure = ctd_deep_data_core$PRESSURE, longitude = ctd_deep_data_core$LONGITUDE, latitude = ctd_deep_data_core$LATITUDE, eos = "gsw")


# Add col for Sigma theta (potential density anomaly)-- input in situ temp values
ctd_deep_data_core$Sigma_theta <- oce::swSigmaTheta(salinity = ctd_deep_data_core$SALINITY, temperature = ctd_deep_data_core$TEMPERATURE, pressure = ctd_deep_data_core$PRESSURE, referencePressure=0, longitude = ctd_deep_data_core$LONGITUDE, latitude = ctd_deep_data_core$LATITUDE, eos = "gsw")

# Add column for spice-- in situ temp: Gibbs SeaWater formulation
ctd_deep_data_core$Spice_gibbs <- oce::swSpice(salinity = ctd_deep_data_core$SALINITY, temperature = ctd_deep_data_core$TEMPERATURE, pressure = ctd_deep_data_core$PRESSURE, longitude = ctd_deep_data_core$LONGITUDE, latitude = ctd_deep_data_core$LATITUDE, eos ="gsw") #Gibbs SeaWater formulation


# Add column for spice-- in situ temp: Flament's formulation
ctd_deep_data_core$Spice_flament <- oce::swSpice(salinity = ctd_deep_data_core$SALINITY, temperature = ctd_deep_data_core$TEMPERATURE, pressure = ctd_deep_data_core$PRESSURE, longitude = ctd_deep_data_core$LONGITUDE, latitude = ctd_deep_data_core$LATITUDE, eos ="unesco") # Flaments
```


# Calculate temperature, oxygen values at 175 m 
NOTE: these data were binned into 2m depth bins, so to calculate values at an odd number, such as 175m which is the upper limit of our meso zone, we have to interpolate from cells above and below
- salinity, temperature, oxygen and density at 175 m (mean between 174 and 176m)
```{r}
ctd_175=ctd_deep_data_core |> 
  filter(CTD_DEPTH==174 | CTD_DEPTH==176) |> # depth above and below
  group_by(year, CTD_INDEX, LATITUDE, LONGITUDE) |> #NOTE CTD_INDEX, LAT LON should be same for each CTD_cast
  summarise(TEMPERATURE=mean(TEMPERATURE),
            SALINITY=mean(SALINITY),
            DENSITY= mean(DENSITY),
            OXYGEN=mean(OXYGEN), 
            .groups = "drop") |> 
            mutate(CTD_DEPTH=175) #add CTD_DEPTH value of 175m

# then add these values to our original dataset
ctd_deep_data_core = rbind(ctd_deep_data_core, ctd_175)


ctd_deep_data_core |> 
  arrange(year, CTD_INDEX, CTD_DEPTH) # check, yes 175 m present
```




# Move to Ocean regrid.Rmd, then satellite.Rmd










