---
title: "Read and visualize output from EV"
output: html_notebook
---

# Objective of code 

This code was written by II to: 
- Load and view processed acoustic data exported from Echoview 
- Calculate echo-integration metrics for each Interval (NASC, Center of Mass, etc.) 
- Plot echograms for each individual day of survey 

Note: These data lead into --> Spatial autocorrelation.Rmd

# Background and Data Description

These data are EK60/EK80 echosounder data collected from the core region of the RREA survey from 2013-2018 and have been filtered, processed, cleaned and exported from Echoview
- Cleaning: impulse noise removal, attenuated signal removal, transient noise and background noise removal
- False bottoms identified via algorithm and removed manually 
- Bad data removal:for each 100m by 5m cell (mean Sv) I set the whole cell to bad data- no data if
there was less than 75% coverage for a given cell (good data-what was left after processing/raw Sv file prior to processing /100 %) 
- These data are from depths greater than 550m deep (bottom depth)
- These data had a -75 dB threshold applied (prior to export, individual sample level (not per grid cell))
- output is mean Sv (dB) data binned by 100m horizontal by 5m vertical cells EK60 data, 38 kHz from 150m to 525m.


# Load libraries
```{r load libraries, echo=FALSE}

library(tidyverse) # all things
library(lubridate) # dates
library(sf) #spatial features

```

# Create list of .csv files to read into R: ALL years of data

These will be read into R in next step
```{r read in EV output}

getwd() # current working directory 
setwd('..') # would like to change the working directory to point to available data stored in parent folder EV_output

# To read 500m_deeper data (100m by 5m output from EV for bottom depths greater than 550m):
# tell directory where files are stored, below is all data in folder for all years
filenames = list.files(path="./EV_output/75percentcoverage_75threshold/Deeper_550m", pattern=, full.names=TRUE)

```

Note: the above only reset the working directory while running the above chunk, and it is reset automatically- can safely ignore warning message

# Read all Echoview output into R

Note this is a lot of data (~4.5 million records), so a little slow

```{r load files into R based on filenames, echo=FALSE}
# remap working directory so R can access output files in parent folder
setwd('..') 

#create df with col for region, year and day of survey- based on filepath name
output <- map_df(filenames, ~read_csv(.x, 
                                      col_types = cols(.default = col_double(), #had to set column type default to avoid conflict error
                                      Time_M=col_guess(), #this is a time value but guess does a good job
                                      Process_ID=col_character()))%>% #except for 2 character cols Time_m and Process_ID
                    mutate(filename=.x))%>% #create a column of filenames for each file
                    mutate(Region=str_sub(filename, -8, -5))%>% #from file name, extract region information
                    mutate(Year=str_sub(filename, -18, -15))%>% # extract year information from file names- new col
                    mutate(Day=str_sub(filename, -13, -10))%>% # extract day information from our file names - new col
                    dplyr::mutate(ID=row_number())%>% #create a col of unique values (row number) in case we need it later
                    #mutate(coverage= (Good_samples/Samples_In_Domain)*100)%>% #percent coverage of a given cell- should all be higher than 75%
                    dplyr::select(-filename)# remove filename col note need to ref package due to conflict with base R

```

NOTES:
- Working directory only changed while running chunk, ignore warning
- Parsing challenges caused the warning message for Process_ID and Time_M columns. However Process_ID is not important and Time_M parses to HMS which is great. for more on parsing:
https://stackoverflow.com/questions/31568409/override-column-types-when-importing-data-using-readrread-csv-when-there-are
- We created a new column for DAY here instead of using Date_M to correspond to the same file "days" we processed for a given "day" in Echoview (naming) given differences in local and GMT time (boat vs recorded time)

# Check range of Sv values:

```{r check range of data values r}
range(output$Sv_mean)

```
The range of values is -999dB to -33.41 for 2013-2018 which is within the range we would expect
From Echoview https://support.echoview.com/WebHelp/Reference/special_data_values.htm
-999dB is the linear equivalent of 0 zero or no backscatter, so should be included! 

# Convert from dB to linear domain
NOTE: all arithmetic needs to be done in the LINEAR domain (sv not Sv) That is why input is Sv, but need to convert via 10^(Sv/10) 
for reference:
```{r convert between linear and dB domain }
# convert from dB to linear
db_to_linear <- function(db){ 10^ (db/10) 
}
# convert from linear to dB
linear_to_db <- function(linear){ 10*log10(linear)
}
# Sv to sv- this doesn't look right! 
Sv_to_sv <- function(Sv){ 10^ (sv/10) 
}
# sv to Sv
sv_to_Sv <- function(sv){ 10*log10(sv)
}
```

### Calculate acoustic metrics from fine resolution (100m horizontal by 5m vertical) bins!
Note: I am only integrating in the vertical direction for now (100m horizontal bins here, but will need to re-bin after spatial autocorrelation analysis)
Note: that for each day of the survey that we processed, there should be unique Interval (horizontal) and layer (vertical) bins.

```{r Acoustic metrics}

#####################################################
#Acoustic Metrics
# definitions from MacLennan et al 2002 and Sam Urmy et al 2012 (vertical metrics)
#####################################################

# Mean Sv - integrated over depths (150m-550m)
Sv_mean_cell= function(Sv_z){
 10*log10(mean(10^(Sv_z/10)))
}

# Cell_H= height of cell (i.e. 5m)
# Sv_Z= Sv at a given depth (ie mean Sv value exported from EV per 100m by 5m cell), at a given depth Depth_mean (mean depth over 5m vertical bin)
# H= sum of all "Height_mean", sum of cell_H


# sa (ABC) Area Backscattering Coefficient- MacLennan et al 2002 

ABC <- function(Sv_z, cell_H){
  sum(10^(Sv_z/10))*cell_H
}
# sum of all sv values--Urmy:depth_integral
# cell_H= height of cell
#note cell_H input is simply a value of 5m (not individual values or you get an error, could also use unique(Height_mean) to generalize it
# Sv_Z= Sv at a given depth (ie mean Sv value exported from EV per 100m by 5m cell), at a given depth Depth_mean (mean depth over 5m vertical bin)
# this is equivalent to the mean sv * total cell thickness (sv_mean*375)


# Calculate NASC for water column 150m-525m
# note this is equivalent to simply taking the sum of all NASC values per cell over our "region" 150-525m
# This is also equal to: our ABC value multiplied by a scaling factor of 4pi (1852^2)- to make the surface area 1nmi squared

NASC <- function(Sv_z, cell_H){
 (sum(10^(Sv_z/10))*cell_H)* 4*pi*1852^2
}

# Sv_Z= Sv at a given depth (ie mean Sv value exported from EV per 100m by 5m cell), at a given depth is the equivalent of our function for ABC * scaling factor
# cell_H= height of cell (5m)


# Calculate NASC from taking the sum of cell nascs
NASC_alt <- function(cell_NASC) {
  sum(cell_NASC)
}


# Calculate Center of Mass (m)- integrated over 150m-550m
CM <- function(z, Sv_z) {
  sum(z*10^(Sv_z/10))/sum(10^(Sv_z/10))
}

# this is an average of depth weighted by sv (volume backscatter) - calculate in sv not Sv space (linear, not dB)
# z is the depth of each of our cells with a given Sv value: input output$Depth_mean
# Sv_Z= Sv at a given depth (ie mean Sv value exported from EV per 100m by 5m cell), at a given depth z


```

# Calculating our acoustics metrics for 100m horizontal bins (2013-2018) 
Note: we calculate the total number of cells per water column here (n), and select for those columns of water with <10 cells missing data \This means at least 65 out of 75 cells (5m vertical cells, 150m-525m). NOTE: whem we later regrid, we omit this step since covering large geographic extent.


```{r calculate acosutic metrics for a given year}

# Calculate acoustic metrics
echo_integration <-  output %>%
                filter(Layer_depth_min>= 150 & # minimum depth of 150m
                      Layer_depth_max<= 525) %>% #max depth of 525m (note this is depth of data, not bottom depth)
               group_by(Year, Day, Interval, Lat_M, Lon_M)%>% #we would like to integrate each 100m horizontal, by 150m-525m water column region
               summarise(
                MVBS= Sv_mean_cell(Sv_z=Sv_mean), #Sv_mean 
                sa= ABC(Sv_z  = Sv_mean, cell_H = 5), # Area backscattering coefficient, ABC, calculated cell-wise
                sA= NASC(Sv_z = Sv_mean, cell_H = 5), #Nautical Area Scattering Coefficient, NASC, calculated cell-wise
                sA_alt= NASC_alt(cell_NASC=NASC), # calc NASC as sum of NASC values
                CM= CM(z=Depth_mean, Sv_z = Sv_mean),
                Sd_Sv= sd(Sv_mean), #calculate the standard deviation of Sv per Interval (diff than EV output?)
                Sd_NASC=sd(NASC), #calculate sd of NASC
                Sd_ABC= sd(ABC), #calculate sd of ABC
                n= n(), # count the number of cells (rows) per Interval-will use this to filter
                .groups = "drop") %>%
              mutate(ABC=10^(MVBS/10)*375)%>% #this is a check on our calculation the mean sv multiplied by total cell height 375m
              mutate(Area_backscattering_strength= linear_to_db(linear=ABC))%>% # log (dB) version of sa (linear)
              mutate(NASC=ABC* 4* pi* 1852^2) %>% #just a calculation check remove when satisfied
              filter(n>= 65) # this selects for those intervals with at least ~87% coverage (allowing 10 cells out of 75 to be bad data, no data)

```


# Echo_integration for ALL data (without 10 cell filter)
I just removed the last line of the code above which removed 100m water columns if more than 10 vertical cells were missing data
This will also be used to calculate the total number of 100m cells per re-grided grid cell! (Regrid.Rmd)
```{r without 10 cell maximum}
echo_integration_all <-  output %>%
                filter(Layer_depth_min>= 150 & # minimum depth of 150m
                      Layer_depth_max<= 525) %>% #max depth of 525m (note this is depth of data, not bottom depth)
               group_by(Year, Day, Interval, Lat_M, Lon_M)%>% #we would like to integrate each 100m horizontal, by 150m-525m water column region
               summarise(
                MVBS= Sv_mean_cell(Sv_z=Sv_mean), #Sv_mean 
                sa= ABC(Sv_z  = Sv_mean, cell_H = 5), # Area backscattering coefficient, ABC, calculated cell-wise
                sA= NASC(Sv_z = Sv_mean, cell_H = 5), #Nautical Area Scattering Coefficient, NASC, calculated cell-wise
                sA_alt= NASC_alt(cell_NASC=NASC), # calc NASC as sum of NASC values
                CM= CM(z=Depth_mean, Sv_z = Sv_mean),
                Sd_Sv= sd(Sv_mean), #calculate the standard deviation of Sv per Interval (diff than EV output?)
                Sd_NASC=sd(NASC), #calculate sd of NASC
                Sd_ABC= sd(ABC), #calculate sd of ABC
                n= n(), # count the number of cells (rows) per Interval-will use this to filter
                .groups = "drop") %>%
              mutate(ABC=10^(MVBS/10)*375)%>% #this is a check on our calculation the mean sv multiplied by total cell height 375m
              mutate(Area_backscattering_strength= linear_to_db(linear=ABC))%>% # log (dB) version of sa (linear)
              mutate(NASC=ABC* 4* pi* 1852^2) 

# to facilitate plotting later, convert Year to year
echo_integration_all = rename(echo_integration_all, year=Year)
```


# FOR LOOP: Create individual echogram of output for each unique day of the survey
This is just for visual reference and quality control check

```{r Echogram of meso depths with acoustic metrics for loop }

# For each year
yr= unique(output$Year) #vector of years

#and day within year, plot an echogram and export file to folder with unique name 

# select for a specific year (2013-2018)
for (i in 1:length(yr)){

    # filter all Sv output for year i  
    yr_echo<- output %>%
            filter(Year %in% yr[i])
  
    # filter all echo-integration results for year i
    yr_int <- echo_integration %>%
            filter(Year %in% yr[i])
  
  #create vector of unique days within a given year
    day_yr= unique(yr_int$Day)
  
  #then for that year, iterate over each day
  for (j in 1:length(day_yr)){
  
  #select Sv output for specific day within a given year
    day_echo <- yr_echo %>% #from year df
                  filter(Day %in% day_yr[j])%>% #select day
                  filter(Layer_depth_min>= 150 & Layer_depth_max<= 525) # select for meso depths

  # select echo-integration output for specific day within a given year 
    day_int <- yr_int %>%
                  filter(Day %in% day_yr[j]) #select day

  # create plot for specific year and day within that year
  ggplot()+
    geom_tile(data= day_echo %>% #Sv df
             filter(Sv_mean > -999)%>% #for dispaly purposes, only show values greater than 0 (-999 dB)
             filter(Interval %in% day_int$Interval), #filer for those intervals that met our criteria for echointegration
            aes(x=Interval, y=Layer_depth_min, fill=Sv_mean))+ #plot Sv per cell echogram background
  scale_fill_viridis_c(direction=1)+ #note c denotes continuous variable, yellows-higher values
  scale_y_reverse()+ #depth shallow to deep
  geom_hline(yintercept=150, color="red", linewidth=1)+ #line for min depth
  geom_hline(yintercept=525, color="red", linewidth=1)+ # line for max depth
  geom_point(data=day_int, aes(x=Interval, y=CM), size=0.5, shape=3)+ #add points for center of mass
  geom_point(data=day_int, aes(x=Interval, y=125, size=NASC, alpha=NASC), color="darkorange3", shape=21)+ #plot NASC lines over exisiting echogram- diff axis?
  ggtitle(label=paste(yr[i], day_yr[j], "Echogram"))+
  labs(x="Interval", y="Depth (m)", fill = "Sv_mean", color="white")+ #legend 
  theme_classic() 
  
  # export plot with unique name
  ggsave(filename=paste0("./echograms/", yr[i], day_yr[j], ".jpeg"))
  }
}
#when done, remove intermediary objects
rm(yr, day_echo, yr_int, day_int, day_yr, yr_echo, echo_integration_day, i, j)#remove iterative objects when all done 
```


# Remove functions and intermideate variables:
leaving until I run spatial autocorrelation analysis 
```{r remove metric calc values}
rm(filenames, ABC, CM, db_to_linear, linear_to_db, NASC, NASC_alt, Sv_mean_cell, sv_to_Sv, Sv_to_sv)
```

# NEXT FILE.....
Head to "spatial autocorrelation.Rmd" to determine the spatial extent at which we do longer have to worry about autocorrelation, then regrid via "regrid.Rmd"



# Summary information
The information below summarizes some simple values about our data: total km traversed, mean depth, etc.

# Total km acoustics data by year
```{r}
#how many total 
echo_integration_all |> 
  group_by(year) |> 
  summarize(no_cells=n(), km=(no_cells*100)/1000) #ea row is a cell, so total number of cells is just total number of rows
  # no_cells: total number of 100m cells
  # km= total kms mutliply the number of cells by 100m, then divide by 1000m to convert from meters to km

# total number of km sampled: 3,442.4 km
687.0	+ 703.3		+ 574.8		+ 598.8		+ 349.1		+529.4 # total of 3,442.4 km sampled 
```
# Depth of acoustics data
```{r}
# download NOAA bathymetry for our region of interest resolution =1 is the highest resolution available
library(marmap) #watch out for overwritting of as.raster!
bathy <- getNOAA.bathy(lon1 =-125 , lon2 = -121, lat1 = 35, lat2 = 39, resolution =  1 )
plot(bathy)
```
# Extract bathy data for acoustic track
```{r}
# can view summary of data
summary.bathy(bathy)

# GET BATHY DATA for ALL collection trips:
bathy_acoustic_track <- get.depth(bathy, x=echo_integration_all$Lon_M, y=echo_integration_all$Lat_M, locator = FALSE, distance=FALSE)

# add depth info to existing dfs
echo_integration_all$depth <- abs(bathy_acoustic_track$depth) #note depths were in negative values so using abs() to convert to positive

```
# Summary of depth info
```{r}
echo_integration_all |> 
  group_by(year) |> 
  summarise(mean_depth=mean(depth), no_days=length(unique(Day)))
  
#overall mean bottom depth
(1535.335+	1492.853+ 1425.323+ 1505.317+ 1533.008+ 1633.758)/6 #1520 m
```
# Plot of acoustic track with depth info
Just for reference
```{r}
library(marmap)
library(ggOceanMaps)

# Plot map of bathymetry data with acoustics overlay
autoplot.bathy(bathy, geom=c("contour", "raster"), show.legend=F)+ #plot as contours and raster (also could use tile but creates white artifact)
scale_fill_etopo()+
#scale_fill_gradient2(low="dodgerblue", mid="gainsboro", high = "darkgreen")+ # color ramp
geom_point(data=echo_integration_all, aes(x=Lon_M, y=Lat_M))+
coord_sf(xlim = c(min(echo_integration_all$Lon_M-0.25), max(echo_integration_all$Lon_M+0.25)), ylim = c(min(echo_integration_all$Lat_M-0.25), max(echo_integration_all$Lat_M+0.25)))+
  facet_wrap(~year)+
  xlab("")+
  ylab("")
```





